### 2017-11-28 paper.tex
Initial Commit, changed language to english

% This template has been tested with LLNCS DOCUMENT CLASS -- version 2.20 (24-JUN-2015)

%"runningheads" enables:
%  - page number on page 2 onwards
%  - title/authors on even/odd pages
%This is good for other readers to enable proper archiving among other papers and pointing to
%content. Even if the title page states the title, when printed and stored in a folder, when
%blindly opening the folder, one could hit not the title page, but an arbitrary page. Therefore,
%it is good to have title printed on the pages, too.
\documentclass[runningheads,a4paper]{llncs}[2015/06/24]

%cmap has to be loaded before any font package (such as cfr-lm)
\usepackage{cmap}
\usepackage[T1]{fontenc}

\usepackage{graphicx}

% für neue deutsche Rechtschreibung
%\usepackage[english,ngerman]{babel}

% für englische Rechtschreibung
%Even though `american`, `english` and `USenglish` are synonyms for babel package (according to https://tex.stackexchange.com/questions/12775/babel-english-american-usenglish), the llncs document class is prepared to avoid the overriding of certain names (such as "Abstract." -> "Abstract" or "Fig." -> "Figure") when using `english`, but not when using the other 2.
%english has to go last to set it as default language
\usepackage[ngerman,english]{babel}

%Eingabeformat UTF-8
\usepackage[utf8]{inputenc}

%Hint by http://tex.stackexchange.com/a/321066/9075 -> enable "= as dashes
\addto\extrasenglish{\languageshorthands{ngerman}\useshorthands{"}}

%cfr-lm is preferred over lmodern. Reasoning at http://tex.stackexchange.com/a/247543/9075
\usepackage[%
rm={oldstyle=false,proportional=true},%
sf={oldstyle=false,proportional=true},%
tt={oldstyle=false,proportional=true,variable=true},%
qt=false%
]{cfr-lm}
%
%if more space is needed, exchange cfr-lm by mathptmx

\graphicspath{{graphics/}}

%Tweaks by IPVS/AS
\usepackage{lncs_as}

%for demonstration purposes only
\usepackage[math]{blindtext}

%Sorts the citations in the brackets
%It also allows \cite{refa, refb}. Otherwise, the document does not compile.
%  Error message: "White space in argument"
\usepackage{cite}

%% If you need packages for other papers,
%% START COPYING HERE
%% COPY ALSO cmap and fontenc from lines 10 to 12

%extended enumerate, such as \begin{compactenum}
\usepackage{paralist}

%put figures inside a text
%\usepackage{picins}
%use
%\piccaptioninside
%\piccaption{...}
%\parpic[r]{\includegraphics ...}
%Text...

%for easy quotations: \enquote{text}
\usepackage{csquotes}

%enable margin kerning
\usepackage{microtype}

%tweak \url{...}
\usepackage{url}
%\urlstyle{same}
%improve wrapping of URLs - hint by http://tex.stackexchange.com/a/10419/9075
\makeatletter
\g@addto@macro{\UrlBreaks}{\UrlOrds}
\makeatother
%nicer // - solution by http://tex.stackexchange.com/a/98470/9075
%DO NOT ACTIVATE -> prevents line breaks
%\makeatletter
%\def\Url@twoslashes{\mathchar`\/\@ifnextchar/{\kern-.2em}{}}
%\g@addto@macro\UrlSpecials{\do\/{\Url@twoslashes}}
%\makeatother

%diagonal lines in a table - http://tex.stackexchange.com/questions/17745/diagonal-lines-in-table-cell
%slashbox is not available in texlive (due to licensing) and also gives bad results. This, we use diagbox
%\usepackage{diagbox}

%required for pdfcomment later
\usepackage[hyperref,svgnames]{xcolor}

\usepackage{listings}
\lstloadlanguages{java}
\lstset{language=java,numbers=left,captionpos=b}

%enable nice comments
%this also loads hyperref
\usepackage{pdfcomment}
%enable hyperref without colors and without bookmarks
\hypersetup{hidelinks,
   colorlinks=true,
   allcolors=black,
   pdfstartview=Fit,
   breaklinks=true}
%enables correct jumping to figures when referencing
\usepackage[all]{hypcap}

\newcommand{\commentontext}[2]{\colorbox{yellow!60}{#1}\pdfcomment[color={0.234 0.867 0.211},hoffset=-6pt,voffset=10pt,opacity=0.5]{#2}}
\newcommand{\commentatside}[1]{\pdfcomment[color={0.045 0.278 0.643},icon=Note]{#1}}

%compatibality with packages todo, easy-todo, todonotes
\newcommand{\todo}[1]{\commentatside{#1}}
%compatiblity with package fixmetodonotes
\newcommand{\TODO}[1]{\commentatside{#1}}

%enable \cref{...} and \Cref{...} instead of \ref: Type of reference included in the link

\usepackage[capitalise,nameinlink,ngerman]{cleveref}
%\usepackage[capitalise,nameinlink,english]{cleveref}
%Nice formats for \cref - only for English texts
%\crefname{section}{Sect.}{Sect.}
%\Crefname{section}{Section}{Sections}

\usepackage{xspace}
%\newcommand{\eg}{e.\,g.\xspace}
%\newcommand{\ie}{i.\,e.\xspace}
\newcommand{\eg}{e.\,g.,\ }
\newcommand{\ie}{i.\,e.,\ }

%introduce \powerset - hint by http://matheplanet.com/matheplanet/nuke/html/viewtopic.php?topic=136492&post_id=997377
\DeclareFontFamily{U}{MnSymbolC}{}
\DeclareSymbolFont{MnSyC}{U}{MnSymbolC}{m}{n}
\DeclareFontShape{U}{MnSymbolC}{m}{n}{
    <-6>  MnSymbolC5
   <6-7>  MnSymbolC6
   <7-8>  MnSymbolC7
   <8-9>  MnSymbolC8
   <9-10> MnSymbolC9
  <10-12> MnSymbolC10
  <12->   MnSymbolC12%
}{}
\DeclareMathSymbol{\powerset}{\mathord}{MnSyC}{180}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

%% END COPYING HERE

\begin{document}

\title{Paper Title}
%If Title is too long, use \titlerunning
%\titlerunning{Short Title}

\author{Vorname Nachname}

\supervisor{Vorname Nachname}

\seminar{Seminarbezeichung}

\semester{WS 2015/2016}

\abgabedatum{Stuttgart, 22.06.2015}

\institute{\email{autor.eins@beispiel.de}}

\frontpagede % creates the frontpage (in German)
%\frontpageen % creates the frontpage (in English)

\thispagestyle{empty}
\cleardoublepage

\maketitle

\begin{abstract}
Die Kurzfassung sollte den Inhalt der Arbeit kurz wiedergeben und aus mindestens 70 und maximal 150 Wörtern bestehen.
Sie wird in einer 9-Punkt Schrift gesetzt und 1 cm eingerückt von rechts und links.
Vor und nach der Kurzfassung sind zwei leere Zeilen.
\end{abstract}

%\begin{keywords}
%keyword1, keyword2
%\end{keywords}

\section{Einleitung}
Ab diesem Abschnitt fängt die eigentliche Arbeit an und man beginnt mit einer Einführung und der Einordnung des Themas.

Absätze werden durch Leerzeilen erreicht.

Als \LaTeX-Editor sind TeXstudio\footnote{\url{http://texstudio.sourceforge.net/}} oder TeXlipse\footnote{\url{http://texlipse.sourceforge.net/}} empfehlenswert.

Für die Versionskontrolle bietet es sich an, jeden Satz in einer neuen Zeile zu beginnen.

Als Nachschlagewerk und gute Einführung in das Thema \LaTeX\ bietet sich \emph{The Not So Short Introduction to \LaTeXe}\ von Oetiker et al an.
Auf der Seite \url{http://tobi.oetiker.ch/lshort/} ist dieses Dokument zum Herunterladen bereitgestellt.

Ein weiteres, empfehlenswertes Dokument ist das \emph{Kochbuch für \LaTeX{}}, welches auf der Seite \url{http://www.uni-giessen.de/hrz/tex/cookbook/cookbook.html} zu finden ist.

Eine Vorlage für Bachelor- und Masterarbeiten ist unter \url{https://github.com/latextemplates/uni-stuttgart-computer-science-template} zu finden.

Für weitere Anleitungen und Informationen empfiehlt sich die Homepage \url{http://www.dante.de/}.
Des Weiteren gibt es eine Reihe von Literatur zum Thema \LaTeX\ von Bürger~\cite{buerger}, Kopka~\cite{kopka}, Günther~\cite{guenther} und vielen anderen.
Weitere Hinweise zu \LaTeX{} unter \url{http://wiki.flupp.de/latex}.
Zu Ausarbeitungen im Allgemeinen unter \url{http://wiki.flupp.de/studium/ausarbeitungen}.

Die Arbeit gleidert sich wie folgt:
...
Schließlich wird in \cref{sec:zusfas} eine Zusammenfassung der Arbeit und ein Ausblick auf weitere Forschungsthemen gegeben.

\subsection{Beispiel für Überschrift 2.~Ebene}

\subsubsection{Beispiel für Überschrift 3.~Ebene}

\paragraph{Beispiel für Überschrift 4.~Ebene}
Dies ist Beispielabsatz zu einem Beispiel einer Überschrift 4.~Ebene.

\section{Formatierungsrichtlinien}

\subsection{Tabellen}
In diesem Abschnitt wird das Einfügen einer Tabelle und das Füllen derselben mit Werten gezeigt.
Im Gegensatz zu Abbildungen erfolgt die Beschriftung üblicherweise oberhalb und nicht unterhalb der Tabelle.
Der letzte Satz der Beschriftung endet ohne Punkt.
Die \Cref{tab:bsp} zeigt die Größenentwicklung der Weltbevölkerung von 8000 v.\,Chr. bis 1980 n.\,Chr.

\begin{table}
\caption{Dieses Beispiel entstammt dem {\it\TeX{}Buch,} S.\,246}
\label{tab:bsp}
\begin{center}
\begin{tabular}{r@{\quad}rl}
\hline
\multicolumn{1}{l}{\rule{0pt}{12pt}
                   Jahr}&\multicolumn{2}{l}{Weltbevölkerung}\\[2pt]
\hline\rule{0pt}{12pt}
8000 v.\,Chr. &     5,000,000& \\
  50 n\,Chr. &   200,000,000& \\
1650 n\,Chr. &   500,000,000& \\
1945 n\,Chr. & 2,300,000,000& \\
1980 n\,Chr. & 4,400,000,000& \\[2pt]
\hline
\end{tabular}
\end{center}
\end{table}

\subsection{Code-Beispiele}

\Cref{lst:example} zeigt ein Code-Beispiel.

\begin{lstlisting}[float,caption=A floating example,label=lst:example]
public static void main(String[] args) {
}
\end{lstlisting}

\subsection{Formeln}
Formeln sind in aufsteigender Reihenfolge, beginnend mit der Nummer~1, durchzunummerieren.
Dies erleichtert den Bezug vom Text auf die zugehörige Formel.
Die Formel
%
\begin{equation}
  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt
\end{equation}
%
berechnet etwas.
Formeln sind bezüglich der Zeichensetzung wie normaler Text zu behandeln, d.\,h.\ falls ein Satz mit einer Formel endet, folgt dieser ein Punkt, der üblicherweise durch ein schmales Leerzeichen abgesetzt wird.

\subsection{Abbildungen}
Im nachfolgenden Beispiel wird dargestellt, wie Grafiken eingefügt werden und wie die Beschriftung bzw.\ die Formatierung derselben auszusehen hat.

Bei Abbildungen erfolgt die Beschriftung unterhalb der Abbildung (siehe \cref{fig:logo}).
Der letzte Satz der Beschreibung endet wie bei Tabellenbeschriftungen ohne Punkt.
Um eine gute Qualität zu erzielen, muss die Auflösung der Abbildungen mindestens 300~DPI betragen.
Bei Strichzeichnungen empfiehlt sich sogar eine weit höhere Auflösung, zwischen 800 und
1200~DPI.

Am besten Vektorgraphiken nehmen und als \texttt{.pdf} exportieren.

\begin{figure}
  \begin{center}
    \includegraphics[width=.5\textwidth]{ipvslogo.png}
    \caption{Das Logo des Instituts}
    \label{fig:logo}
   \end{center}
\end{figure}

\subsection{Literaturverweise}
Ein Verweis auf eine im Literaturverzeichnis aufgeführte
Literaturquelle wird durch die in eckige Klammern eingeschlossene
Nummer der entsprechenden Literaturquelle angegeben. Als Beispiel ist
hier nochmals auf Buerger~\cite{buerger} verwiesen.

\subsection{Anführungszeichen}
Entweder "`so"' tippen oder \enquote{so}.
``So'' bitte nicht, das führt zu englischen Anführungszeichen.

\subsection{Sonstiges}
Brackets work as designed:
<test>

The symbol for powerset is now correct: $\powerset$ and not a Weierstrass p ($\wp$).

\begin{inparaenum}
\item All these items...
\item ...appear in one line
\item This is enabled by the paralist package.
\end{inparaenum}

``something in quotes'' using plain tex or use \enquote{the enquote command}.

You can now write words containing hyphens which are hyphenated (application"=specific) at other places.
This is enabled by an additional configuration of the babel package.
In case you write \enquote{application-specific}, then the word will only be hyphenated at the dash.
You can also write applica\allowbreak{}tion-specific, but this is much more effort.

\section{Zusammenfassung}
\label{sec:zusfas}
Dieser Absatz schließt die Arbeit ab.

\subsubsection*{Acknowledgments}
...

In the bibliography, use \texttt{\textbackslash textsuperscript} for ``st'', ``nd'', ...:
E.g., \enquote{The 2\textsuperscript{nd} conference on examples}.
When you use \href{https://www.jabref.org}{JabRef}, you can use the clean up command to achieve that.
See \url{https://help.jabref.org/en/CleanupEntries} for an overview of the cleanup functionality.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{splncs03}
\bibliography{paper}
Alle Links wurden zuletzt am 22.06.2015 geprüft.
\end{document}

### 2017-11-28 paper.tex
Minor Introduction done

@@
 
-\usepackage[capitalise,nameinlink,ngerman]{cleveref}
-%\usepackage[capitalise,nameinlink,english]{cleveref}
+%\usepackage[capitalise,nameinlink,ngerman]{cleveref}
+\usepackage[capitalise,nameinlink,english]{cleveref}
 %Nice formats for \cref - only for English texts
@@
 
-\title{Paper Title}
+\title{Apache Storm}
 %If Title is too long, use \titlerunning
@@
 
-\author{Vorname Nachname}
+\author{Aanal Raj Basaula}
 
-\supervisor{Vorname Nachname}
+\supervisor{Matthias Wieland}
 
@@
 
-\semester{WS 2015/2016}
+\semester{WS 2017/2018}
 
-\abgabedatum{Stuttgart, 22.06.2015}
+\abgabedatum{Stuttgart, 11.12.2017}
 
-\institute{\email{autor.eins@beispiel.de}}
+\institute{\email{aanal.basaula@gmail.com}}
 
-\frontpagede % creates the frontpage (in German)
-%\frontpageen % creates the frontpage (in English)
+%\frontpagede % creates the frontpage (in German)
+\frontpageen % creates the frontpage (in English)
 
@@
 \begin{abstract}
-Die Kurzfassung sollte den Inhalt der Arbeit kurz wiedergeben und aus mindestens 70 und maximal 150 Wörtern bestehen.
-Sie wird in einer 9-Punkt Schrift gesetzt und 1 cm eingerückt von rechts und links.
-Vor und nach der Kurzfassung sind zwei leere Zeilen.
+In today's world everyone is connected to this ever increasing network of computers be it via a Personal computer, a smart phone or an Embedded Device such as your smart home appliances.
+These devices generate a huge amounts of data every instance, that can be analyzed to leverage business opportunities previously undiscovered.
+To be able to discern these windows of opportunities, one must be able to process and analyze the data.
+In this report we talk about such scenario and the available tools which allows us to process large data streams with minimum latency.
 \end{abstract}
 
-%\begin{keywords}
-%keyword1, keyword2
-%\end{keywords}
+\begin{keywords}
+Apache Storm, Stream Processing, Real Time Processing, Big Data
+\end{keywords}
+
+\section{Introduction}
+
+\textbf{Data} today is generated by a lot of devices that are connected to the internet and the amount of data being generated at any instance is increasing rapidly.
+The above mentioned device could be a Personal Computer, a Smart phone, or even an Embedded System.
+Due to the rapid adoption of Internet of Things, data is being generated at a rate that is unprecedented.
+The sheer volume of data generated and at the rate at which it is being generated can be termed as Big Data.
+\textbf{Big Data} made it's appearance in 1998 SGI slide by John Mashey and now is of interest to a lot of financial as well as research institutions.[1]
+
+Data can be considered the oil of the current age, because with huge amounts of data one can determine hidden patterns.
+A lot of companies utilize pattern recognition on big data to open opportunities for revenue generation whereas research institutions use it to discover something new.
+The question on how to classify any data as big data can be answered by using the 3Vs of big data: Volume, Variety and Velocity.
+Any data with big volume or variety (such as text, video, images, etc) or that gets generated rapidly can be considered as big data.
+In this paper regarding \textbf{Apache Storm} we consider the third V of big data to be the influencing factor.
+
+\textbf{Data Stream} is a sequence of data flowing from a source to a destination, for example a sensor reporting the temperature to a server.
+The sensor measures the temperature periodically and updates it to the server.
+Another example could also be the purchase requests of your product.
+In all of these scenarios, the data is continuously being generated.
+Traditionally analyzing of these types of data requires storage of these data onto persistence and later batch processing it to discern patterns as well as other useful information.
+But, with the help of \textbf{Apache Storm} we are able to process these data as they are generated with minimum amount of latency.
+
+This benefit could be a major game changer for many companies as Stream processing of data allows Real Time Analysis.
+As an example, an online store could monitor its purchase requests for best performing and least performing articles and could offer discounts on the articles not performing good to increase their sales volume.
+This small change could increase the earnings of a company and will surely be advantageous.
+
+\section{Apache Storm: Concepts}
 

### 2017-11-30 paper.tex
Added Concepts of Apache Storm

@@
 A lot of companies utilize pattern recognition on big data to open opportunities for revenue generation whereas research institutions use it to discover something new.
-The question on how to classify any data as big data can be answered by using the 3Vs of big data: Volume, Variety and Velocity.
+The question on how to classify any data as big data can be answered by using the 3Vs of big data: Volume, Variety and Velocity [2].
 Any data with big volume or variety (such as text, video, images, etc) or that gets generated rapidly can be considered as big data.
@@
 \section{Apache Storm: Concepts}
+
+Apache storm is one of the available open source tools for stream data processing, which is considered to be one of the most easy to use, fast and fault tolerant systems available [3].
+These three factors have been the major points for the general public to choose \textbf{Apache Storm} above other available tools.
+It was originally created by Nathan Marz at BackType and till now, at present more than 60 companies use or experiment with this tool.
+
+\subsection{Architecture}
+The basic Storm data processing architecture consists of stream of tuples flowing through a defined topology to yield the required analytic result.
+Processing of this stream data is performed in Nodes called the worker nodes, which are responsible for performing the required calculations.
+There is also another node named the Master Node, which is responsible for the complete Storm infrastructure,  by keeping track of all the worker nodes and the total topology itself.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=.5\textwidth]{arch.png}
+    \caption{High Level Storm Architecture}
+    \label{fig:arch}
+   \end{center}
+\end{figure}
+
+\subsubsection{Nimbus}
+
+\paragraph{The Master node} also called the \textbf{Nimbus} is also responsible for distribution and coordination of the execution of topologies.
+The Nimbus is the first point where the user submits the defined topology, which is then distributed among the worker nodes.
+Figure \ref{fig:arch} illustrates the high level architecture where, Nimbus monitors all the available worker nodes and both the nimbus as well as the worker nodes use \textbf{Apache Zookeeper} to store their states.
+The actual processing is done by the \textbf{worker nodes}, and each worker node runs multiple Worker Processes each of them mapped to their own topology [4].
+It is also worth while noting that multiple worker processes in a worker node could be executing different parts of the topology.
+
+\subsubsection{Worker Node}
+\paragraph{Worker Node} is responsible for all the heavy lifting required for processing of the topology.
+One worker node may run a multiple number of worker processes.
+Each of these worker processes may belong to the same topology or different topologies.
+A worker node maintains a \textbf{Supervisor} and a multiple number of Worker Processes which in turn maintains a multiple number of executors.
+Thus, Executors enable intra-topology parallelism whereas Tasks provide intra-spout or intra-bolt parallelism.
+Figure \ref{fig:workerarch} depicts the design of a Worker node.[4]
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=.5\textwidth]{worker.png}
+    \caption{Worker Node Architecture}
+    \label{fig:workerarch}
+   \end{center}
+\end{figure}
+
+\paragraph{Supervisor} is running in every Worker Node and communicates with the Nimbus to receive assignments from it.
+Apart from taking assignments from the Nimbus it is also responsible for monitoring the health of the of the worker processes and respawns the worker processes in case it dies.
+
+\paragraph{Worker Process} is the Entity that does the processing and is mapped to a certain topology.
+Each worker node spawns a JVM and runs one or multiple executors.
+These executors may be running any of the bolt or spout within the topology, thus providing an added level of parallelism.
+
+\paragraph{Executors} are made up of one or multiple Tasks.
+Unless specified explicitly, storm assigns single task to a single executor.
+Executors takes tuples from the in queue, examines the task this tuple is destined to, performs the task and then places the result in the out queue [4].
+
+\subsubsection{Topology}
+\paragraph{Topology} is a directed graph where the vertices represent computation and the edges represents the data flow.
 

### 2017-12-01 paper.tex
Added Topology info

@@
 \paragraph{Topology} is a directed graph where the vertices represent computation and the edges represents the data flow.
+Another important thing about topologies in Storm is that they are allowed to have cycles.
+Figure \ref{fig:topo} shows a basic Storm Topology, which consists of Spouts and Bolts.
+Spouts are the Sources of tuples in a Topology whereas Bolts are the units which process these generated tuples.
+The user first submits the topology to the Apache Storm Nimbus, which consults with each of the available Supervisors in the worker nodes and assigns each of them with tasks related to the topology.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=.7\textwidth]{topo.png}
+    \caption{Generic Storm Topology}
+    \label{fig:topo}
+   \end{center}
+\end{figure}
+
+\paragraph{Spout} is the entry point for any data in the Storm Topology.
+As shown in the figure \ref{fig:topo} spouts take data from external source and emits them into the topology as tuples.
+The input source could be a Heterogenous data source.
+
+\paragraph{Bolt} are the processing entities of a Topology.
+It receives raw tuples emitted by spouts or processed tuples sent by another bolt and sends it down the pipe after it has completed its processing.
 

### 2017-12-08 paper.tex
Started with Practical Use case

@@
 
-\subsubsection{Topology}
+\subsection{Topology}
 \paragraph{Topology} is a directed graph where the vertices represent computation and the edges represents the data flow.
@@
 
-\paragraph{Spout} is the entry point for any data in the Storm Topology.
+\subsubsection{Spout} is the entry point for any data in the Storm Topology.
 As shown in the figure \ref{fig:topo} spouts take data from external source and emits them into the topology as tuples.
@@
 
-\paragraph{Bolt} are the processing entities of a Topology.
+\subsubsection{Bolt} are the processing entities of a Topology.
 It receives raw tuples emitted by spouts or processed tuples sent by another bolt and sends it down the pipe after it has completed its processing.
+
+\section{Practical Use Case}
+Apache Storm is used for stream processing of data, in many sources [2,3] a word frequency count example has been shown which can be considered to be trivial, but in actual practice it can be used to execute complicated algorithms.
+For the purposes of this paper we consider a trivial but different example of \textbf{Sentiment Analysis}.
+
+\subsection{Scenario}
+
+\subsubsection{Description}
+\paragraph{Let us consider} a company which produces a range of products, for example \textbf{Apple Inc.} produces IPhones, IPods, MacBooks, etc.
+For such a big company an understanding of the sentiment of the people is important to maintain a positive outlook.
+The company can also analyze the feedback people place on a specific product, to change and adjust the pricing schemes in real time to the demand.
+At present there are many social media networks available and people tend to share their views on these networks,  \textbf{Twitter} is also one of these social networks where a number of posts are available publicly and can be analyzed to obtain a generalized view about a specific product or the company in general.
+
+\subsubsection{Proposed Solution}
+
+\paragraph{Solution} to the scenario can be implemented using \textbf{Apache Storm} to process the tweets in real time.
+Figure \ref{fig:solution} shows the simplified system overview, consisting of Twitter Streaming Servers as the source of information.
+This information is processed in the Storm Clusters from within the Company Infrastructure and the analyzed result is reported to the User Interface, which can be accessed by the appropriate user.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=.7\textwidth]{solution.png}
+    \caption{Generic Streaming Solution implemented in this use case}
+    \label{fig:solution}
+   \end{center}
+\end{figure}
+
+\paragraph{This} is not the only possible solution, there are multiple other tools such as \textbf{Spark Streaming}, \textbf{Samza}, etc.
+but this paper considers Apache Storm as it is in the spotlight of this paper, for further details regarding other available tools and their comparisons with Apache Storm, please consider the section \ref{sec:othertools} for further details.
+
+\subsection{Practical Implementation}
+
+\section{Other Tools}
+ \label{sec:othertools}
 

### 2017-12-10 paper.tex
Base Outline done for submission

@@
 In today's world everyone is connected to this ever increasing network of computers be it via a Personal computer, a smart phone or an Embedded Device such as your smart home appliances.
-These devices generate a huge amounts of data every instance, that can be analyzed to leverage business opportunities previously undiscovered.
-To be able to discern these windows of opportunities, one must be able to process and analyze the data.
+These devices generate huge amounts of data every instance, that can be analyzed to uncover patterns, some of which can be used to improve businesses as well.
 In this report we talk about such scenario and the available tools which allows us to process large data streams with minimum latency.
+Specifically Apache Storm, a publicly available Stream Processing Tool.
+We will dive into the concepts of Storm, how one can design usable systems with it and also compare storm to the other available tools.
 \end{abstract}
@@
 \textbf{Data} today is generated by a lot of devices that are connected to the internet and the amount of data being generated at any instance is increasing rapidly.
-The above mentioned device could be a Personal Computer, a Smart phone, or even an Embedded System.
-Due to the rapid adoption of Internet of Things, data is being generated at a rate that is unprecedented.
-The sheer volume of data generated and at the rate at which it is being generated can be termed as Big Data.
+These devices could include a personal computer, a smart phone, or even an embedded device.
+Due to the rapid adoption of Internet of Things, data is being generated at an unprecedented rate.
+The sheer volume of data generated and the rate at which it is being generated can be termed as Big Data.
 \textbf{Big Data} made it's appearance in 1998 SGI slide by John Mashey and now is of interest to a lot of financial as well as research institutions.[1]
 
-Data can be considered the oil of the current age, because with huge amounts of data one can determine hidden patterns.
-A lot of companies utilize pattern recognition on big data to open opportunities for revenue generation whereas research institutions use it to discover something new.
-The question on how to classify any data as big data can be answered by using the 3Vs of big data: Volume, Variety and Velocity [2].
+Data can be considered the oil of this present age, because with huge amounts of data one can determine hidden patterns.
+A lot of companies utilize pattern recognition on big data to open opportunities for revenue generation whereas research institutions use it to discover patterns simply not observable by the human mind(simply due to the volume of data).
+The question on how to classify any data as big data can be answered by using the 3Vs: Volume, Variety and Velocity [2].
 Any data with big volume or variety (such as text, video, images, etc) or that gets generated rapidly can be considered as big data.
@@
 The sensor measures the temperature periodically and updates it to the server.
-Another example could also be the purchase requests of your product.
+Another example could also be the purchase requests of products in a shopping site.
 In all of these scenarios, the data is continuously being generated.
@@
 As an example, an online store could monitor its purchase requests for best performing and least performing articles and could offer discounts on the articles not performing good to increase their sales volume.
-This small change could increase the earnings of a company and will surely be advantageous.
+This small change could increase the gross income of a company and will surely be advantageous.
 
 \section{Apache Storm: Concepts}
+\label{sec:concepts}
 
@@
   \begin{center}
-    \includegraphics[width=.5\textwidth]{arch.png}
+    \includegraphics[width=.7\textwidth]{arch.png}
     \caption{High Level Storm Architecture}
@@
   \begin{center}
-    \includegraphics[width=.5\textwidth]{worker.png}
+    \includegraphics[width=.7\textwidth]{worker.png}
     \caption{Worker Node Architecture}
@@
 
-\subsubsection{Spout} is the entry point for any data in the Storm Topology.
+\subsubsection{Spout} 
+\paragraph{Spouts} are the entry point for any data in the Storm Topology.
 As shown in the figure \ref{fig:topo} spouts take data from external source and emits them into the topology as tuples.
@@
 
-\subsubsection{Bolt} are the processing entities of a Topology.
+\subsubsection{Bolt} 
+\paragraph{Bolts} are the processing entities of a Topology.
 It receives raw tuples emitted by spouts or processed tuples sent by another bolt and sends it down the pipe after it has completed its processing.
+
+\section{System Design}
+Storm server clusters as discussed above in Section \ref{sec:concepts}, requires at least a Master node and a series of Worker nodes.
+Apart from these, it also requires zookeeper to keep track of the status of all the master as well as the worker nodes.
+Setup of these nodes are trivial and can be configured easily in storm using the configuration files.
+But, the design of a system with regard to the requirements is not such a trivial task because of the different data processing needs.
+The things to consider while designing a Storm Solution is the System architecture i.e.
+Lambda or Lambda-less.
+After selecting one of the above solution, further decisions for the topology design should be considered.
+
+\subsection{System Architecture}
+Apache storm is a flexible software which is capable of adapting to the requirements.
+It can be used in a lambda architecture or a lambda less architecture.Depending upon the requirements of the system, one can opt for either of the above solutions
+
+\subsubsection{Lambda-less Architecture}
+\paragraph{Apache Storm} can handle data processing without the need of persistence layer and offers its own 
+resiliency mechanisms to make sure data is tuples injected into the system are processed.
+It has two types of processing schemes, \textbf{at least once} where each tuple is processed as the name suggests, at least once, and the other \textbf{at most once}, where the tuples are either processed once or dropped in cases of failure.
+To provide \textbf{at least once} semantics, Storm maintains an acker bolt, which keeps track of the tuple flow in the directed acyclic graph.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=.7\textwidth]{acker.png}
+    \caption{Storm Acker Bolt [4]}
+    \label{fig:acker}
+   \end{center}
+\end{figure}
+
+Figure \ref{fig:acker} shows the Acker bolt integration with the storm topology.
+It keeps track of the tuple flow starting from the point where it has been injected into the topology till it leaves the topology.
+After the tuple has left the topology, an \textbf{ack} is sent to the Spout, at this point, it can decide to remove the Tuple from memory.
+
+\subsubsection{Lambda Architecture}
+\paragraph{Lambda Architecture} is a data processing Architecture designed to handle high volume data utilizing both batch and stream processing.
+Apache storm allows users to integrate Persistence models into the system.
+This flexibility allows users to integrate batch processing into the system as well as seen in figure \ref{fig:lambda}.
+
+Data is persisted in the persistence layer so that in cases of high data volume, Batch processing can be done on the input data to provide guaranteed processing and to reduce the computation requirement on the Real Time Data Processing Unit.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=.7\textwidth]{lambda.png}
+    \caption{Lambda Architecture}
+    \label{fig:lambda}
+   \end{center}
+\end{figure}
+
+\subsection{Topology Design Considerations}
+A storm topology design with the spouts and bolts should be properly considered.
+A spout should inject data into the topology from external sources where as the bolt should be a smallest implementable task.
+A storm topology should also be considered with various considerations in mind:
+
+\subsubsection{Scalability:} A designed topology should be considered with scalability in mind.
+The major question is what configuration of bolts enables us to process greater amount of data at real time.
+Let us consider a simple example of word count topology which counts the repetition of words and ranks the most used words.
+It can be performed by simply including a spout and two bolts, the first of which to count the number of words and the second to aggregate and rank it (see fig.
+\ref{fig:ranking} Topology Setup 1).
+Since we should have only one instance of Aggregator, to be able to get the aggregated result, in cases of increased data load, the number of rolling count bolts can be increased, but the Aggregation Bolt cannot be increases thus adding increased load at this node.
+To relax the demand on this node, we can introduce another bolt, the Intermediate Ranking Bolt, which is responsible for ranking the words at a certain predefined intervals and passes the intermediate ranks to the aggregator.
+This reduces load on the Aggregator since we can increase the intermediate ranking bolts to cope with the increased load.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{ranking.png}
+    \caption{Possible Topologies for Word Count and Ranking}
+    \label{fig:ranking}
+   \end{center}
+\end{figure}
+\subsubsection{Resilience:} Apache storm includes APIs which allow users to develop solutions which are resilient in nature.
+Using the acker bolt as show in figure \ref{fig:acker}, users can track the progress of the tuple injected into the system.
+Thus, writing a spout which can replay the tuple in case of failures allows us to guarantee the processing of data.
+\subsubsection{Interoperability:} The designed topology should be interoperable with different sources of data, and by default should work out of the box by adding simply other Spouts in the topology.
+This also amounts to the maintainability of the code.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{ranking.png}
+    \caption{Example Ranking Topologies}
+    \label{fig:ranking}
+   \end{center}
+\end{figure}
 
@@
 \subsection{Practical Implementation}
+The implementation of this solution concluded in three sub phases, namely Design, Development and Deployment.
+In the design phase, the storm topology was thoroughly planned, specifically the spouts, bolts and how they are interconnected.
+In the development phase, these spouts, bolts and other required aspects of the software was programmed and lastly was deployed on a locally hosted storm server.
+
+\subsubsection{Design}
+\paragraph{Designing} a storm topology consists of determining the external sources of data (spouts), the processing that is required to this data imported from external sources (bolts) and the target system.
+Refer to figure \ref{fig:topoimpl} for the specific topology implementation for our use case.
+The sources of information for the scope of this paper is only Twitter Streaming API therefore a single Spout for integrating with the Twitter API is sufficient.
+
+To Process the stream of tweets, we need to parse these tweets injected into the topology.
+The ParseTweetBolt processes these tweets and then emits whether the tweet was positive or negative under the simplified scenario.
+These separate positive or negative remark on a tweet does not provide useful information unless it is aggregated, thus a ResultAggregationBolt is responsible for aggregating all the results of the Parser bolt and providing the result to the target in real time.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=.7\textwidth]{topoimpl.png}
+    \caption{Topology Implementation for Sentiment Analysis}
+    \label{fig:topoimpl}
+   \end{center}
+\end{figure}
+
+The tweet spout contacts the twitter streaming server and fetches the available tweets in real time and relays it into the topology as a single tweet per tuple.
+The tuple is forwarded to the TweetParseBolt which will logically analyze the tweet as a positive or a negative tweet and with the percentage probability and emits this result to the ResultAggregationBolt.
+The ResultAggregationBolt calculates simply the weighted mean of the positive feedback and the negative feedbacks and outputs it into a local file, which can be later observed.
+
+\subsubsection{Development}
+\paragraph{Writing} code for storm can be achieved by using different programming languages depending upon the user.
+Apache Storm was designed to be usable with any programming languages\footnote{\url{http://storm.apache.org/about/multi-language.html}} and currently has Adapters available for Ruby, Python, Javascript and Perl.
+
+The only spout uses Twitter Hosebird Client (\url{https://github.com/twitter/hbc}) to connect to Twitter Streaming Server and consume the Tweets generated in real time.
+Whereas the Bolts were programmed in java and the code can be viewed in \url{https://github.com/hakku276/storm-sentiment-analysis}.
 
@@
  \label{sec:othertools}
+
+ <Describe shortly that other tools are available as well and list them with short descriptions>
+
+ \subsection{Comparision between Storm, Spark and Samza}
+
+ <Gather and show the differences between Storm, Spark and Samza in general>
+
+ \subsubsection{Coding}
+ <Compare how the coding is in all these platforms>
+
+ \subsubsection{Tool setup}
+ <Compare how easy it is to setup the tools>
+
+ \subsubsection{Deployment}
+ <Compare how easy it is to deploy the code onto the tool>
+
+ \subsubsection{Resilience}
+ <Compare the resilience model of the tools>
+
+ \section{Discussion}
+<What I felt about the tool, ease of use and other general aspects>
+
+ The zookeeper services can be run in a single machine or can be deployed in a cluster as well.
+A single zookeeper instance can also be shared between other services and Apache Storm, but, the major consideration to be taken into account is performance.
+If there are multiple number of systems that utilize the zookeeper, the Zookeeper can run out of number of clients it can service, thus affecting up-time of the systems in general.[4] Whereas having a dedicated Zookeeper allows a high performing cluster.
 
@@
 
-Ein weiteres, empfehlenswertes Dokument ist das \emph{Kochbuch für \LaTeX{}}, welches auf der Seite \url{http://www.uni-giessen.de/hrz/tex/cookbook/cookbook.html} zu finden ist.
+Ein weiteres, empfehlenswertes Dokument ist das \emph{Kochbuch für \LaTeX{}}, welches auf der Seite \url{http://www.uni-giessen.de/hrz/tex/cookbook/cookbook.html} zu finden ist.``````````````2
 

### 2017-12-10 paper.tex
Changed Abstract

@@
 
-\seminar{Seminarbezeichung}
+\seminar{Advanced Topics in Data Management}
 
@@
 
-\institute{\email{aanal.basaula@gmail.com}}
+\institute{\email{st152995@stud.uni-stuttgart.de}}
 
@@
 \begin{abstract}
-In today's world everyone is connected to this ever increasing network of computers be it via a Personal computer, a smart phone or an Embedded Device such as your smart home appliances.
-These devices generate huge amounts of data every instance, that can be analyzed to uncover patterns, some of which can be used to improve businesses as well.
-In this report we talk about such scenario and the available tools which allows us to process large data streams with minimum latency.
-Specifically Apache Storm, a publicly available Stream Processing Tool.
-We will dive into the concepts of Storm, how one can design usable systems with it and also compare storm to the other available tools.
+Apache Storm is one of the available open source tools for Stream Data Processing.
+This document provides a brief overview of the intricacies of the Architecture of Storm and tends to proceed further on the details of considerations required to design a Solution based on Storm.
+The design considerations relating to deployment of Storm System as a Lambda or a Lambda-less Architecture and the sketch of the various internal components of Storm i.e.
+Bolts, Spouts.
+We consider a practical implementation of Storm to try and express the theoretical design considerations in real practice.
+Finally we list out the other available tools and compare them to Storm with regard to Performance, Resilience and various other parameters.
 \end{abstract}
@@
 
- The zookeeper services can be run in a single machine or can be deployed in a cluster as well.
-A single zookeeper instance can also be shared between other services and Apache Storm, but, the major consideration to be taken into account is performance.
-If there are multiple number of systems that utilize the zookeeper, the Zookeeper can run out of number of clients it can service, thus affecting up-time of the systems in general.[4] Whereas having a dedicated Zookeeper allows a high performing cluster.
-
-\section{Einleitung}
-Ab diesem Abschnitt fängt die eigentliche Arbeit an und man beginnt mit einer Einführung und der Einordnung des Themas.
-
-Absätze werden durch Leerzeilen erreicht.
-
-Als \LaTeX-Editor sind TeXstudio\footnote{\url{http://texstudio.sourceforge.net/}} oder TeXlipse\footnote{\url{http://texlipse.sourceforge.net/}} empfehlenswert.
-
-Für die Versionskontrolle bietet es sich an, jeden Satz in einer neuen Zeile zu beginnen.
-
-Als Nachschlagewerk und gute Einführung in das Thema \LaTeX\ bietet sich \emph{The Not So Short Introduction to \LaTeXe}\ von Oetiker et al an.
-Auf der Seite \url{http://tobi.oetiker.ch/lshort/} ist dieses Dokument zum Herunterladen bereitgestellt.
-
-Ein weiteres, empfehlenswertes Dokument ist das \emph{Kochbuch für \LaTeX{}}, welches auf der Seite \url{http://www.uni-giessen.de/hrz/tex/cookbook/cookbook.html} zu finden ist.``````````````2
-
-Eine Vorlage für Bachelor- und Masterarbeiten ist unter \url{https://github.com/latextemplates/uni-stuttgart-computer-science-template} zu finden.
-
-Für weitere Anleitungen und Informationen empfiehlt sich die Homepage \url{http://www.dante.de/}.
-Des Weiteren gibt es eine Reihe von Literatur zum Thema \LaTeX\ von Bürger~\cite{buerger}, Kopka~\cite{kopka}, Günther~\cite{guenther} und vielen anderen.
-Weitere Hinweise zu \LaTeX{} unter \url{http://wiki.flupp.de/latex}.
-Zu Ausarbeitungen im Allgemeinen unter \url{http://wiki.flupp.de/studium/ausarbeitungen}.
-
-Die Arbeit gleidert sich wie folgt:
-...
-Schließlich wird in \cref{sec:zusfas} eine Zusammenfassung der Arbeit und ein Ausblick auf weitere Forschungsthemen gegeben.
-
-\subsection{Beispiel für Überschrift 2.~Ebene}
-
-\subsubsection{Beispiel für Überschrift 3.~Ebene}
-
-\paragraph{Beispiel für Überschrift 4.~Ebene}
-Dies ist Beispielabsatz zu einem Beispiel einer Überschrift 4.~Ebene.
-
-\section{Formatierungsrichtlinien}
-
-\subsection{Tabellen}
-In diesem Abschnitt wird das Einfügen einer Tabelle und das Füllen derselben mit Werten gezeigt.
-Im Gegensatz zu Abbildungen erfolgt die Beschriftung üblicherweise oberhalb und nicht unterhalb der Tabelle.
-Der letzte Satz der Beschriftung endet ohne Punkt.
-Die \Cref{tab:bsp} zeigt die Größenentwicklung der Weltbevölkerung von 8000 v.\,Chr. bis 1980 n.\,Chr.
-
-\begin{table}
-\caption{Dieses Beispiel entstammt dem {\it\TeX{}Buch,} S.\,246}
-\label{tab:bsp}
-\begin{center}
-\begin{tabular}{r@{\quad}rl}
-\hline
-\multicolumn{1}{l}{\rule{0pt}{12pt}
-                   Jahr}&\multicolumn{2}{l}{Weltbevölkerung}\\[2pt]
-\hline\rule{0pt}{12pt}
-8000 v.\,Chr. &     5,000,000& \\
-  50 n\,Chr. &   200,000,000& \\
-1650 n\,Chr. &   500,000,000& \\
-1945 n\,Chr. & 2,300,000,000& \\
-1980 n\,Chr. & 4,400,000,000& \\[2pt]
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\subsection{Code-Beispiele}
-
-\Cref{lst:example} zeigt ein Code-Beispiel.
-
-\begin{lstlisting}[float,caption=A floating example,label=lst:example]
-public static void main(String[] args) {
-}
-\end{lstlisting}
-
-\subsection{Formeln}
-Formeln sind in aufsteigender Reihenfolge, beginnend mit der Nummer~1, durchzunummerieren.
-Dies erleichtert den Bezug vom Text auf die zugehörige Formel.
-Die Formel
-%
-\begin{equation}
-  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
-  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt
-\end{equation}
-%
-berechnet etwas.
-Formeln sind bezüglich der Zeichensetzung wie normaler Text zu behandeln, d.\,h.\ falls ein Satz mit einer Formel endet, folgt dieser ein Punkt, der üblicherweise durch ein schmales Leerzeichen abgesetzt wird.
-
-\subsection{Abbildungen}
-Im nachfolgenden Beispiel wird dargestellt, wie Grafiken eingefügt werden und wie die Beschriftung bzw.\ die Formatierung derselben auszusehen hat.
-
-Bei Abbildungen erfolgt die Beschriftung unterhalb der Abbildung (siehe \cref{fig:logo}).
-Der letzte Satz der Beschreibung endet wie bei Tabellenbeschriftungen ohne Punkt.
-Um eine gute Qualität zu erzielen, muss die Auflösung der Abbildungen mindestens 300~DPI betragen.
-Bei Strichzeichnungen empfiehlt sich sogar eine weit höhere Auflösung, zwischen 800 und
-1200~DPI.
-
-Am besten Vektorgraphiken nehmen und als \texttt{.pdf} exportieren.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=.5\textwidth]{ipvslogo.png}
-    \caption{Das Logo des Instituts}
-    \label{fig:logo}
-   \end{center}
-\end{figure}
-
-\subsection{Literaturverweise}
-Ein Verweis auf eine im Literaturverzeichnis aufgeführte
-Literaturquelle wird durch die in eckige Klammern eingeschlossene
-Nummer der entsprechenden Literaturquelle angegeben. Als Beispiel ist
-hier nochmals auf Buerger~\cite{buerger} verwiesen.
-
-\subsection{Anführungszeichen}
-Entweder "`so"' tippen oder \enquote{so}.
-``So'' bitte nicht, das führt zu englischen Anführungszeichen.
-
-\subsection{Sonstiges}
-Brackets work as designed:
-<test>
-
-The symbol for powerset is now correct: $\powerset$ and not a Weierstrass p ($\wp$).
-
-\begin{inparaenum}
-\item All these items...
-\item ...appear in one line
-\item This is enabled by the paralist package.
-\end{inparaenum}
-
-``something in quotes'' using plain tex or use \enquote{the enquote command}.
-
-You can now write words containing hyphens which are hyphenated (application"=specific) at other places.
-This is enabled by an additional configuration of the babel package.
-In case you write \enquote{application-specific}, then the word will only be hyphenated at the dash.
-You can also write applica\allowbreak{}tion-specific, but this is much more effort.
-
-\section{Zusammenfassung}
-\label{sec:zusfas}
-Dieser Absatz schließt die Arbeit ab.
-
-\subsubsection*{Acknowledgments}
-...
-
-In the bibliography, use \texttt{\textbackslash textsuperscript} for ``st'', ``nd'', ...:
-E.g., \enquote{The 2\textsuperscript{nd} conference on examples}.
-When you use \href{https://www.jabref.org}{JabRef}, you can use the clean up command to achieve that.
-See \url{https://help.jabref.org/en/CleanupEntries} for an overview of the cleanup functionality.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\bibliographystyle{splncs03}
-\bibliography{paper}
-Alle Links wurden zuletzt am 22.06.2015 geprüft.
 \end{document}

### 2017-12-15 paper.tex
Checkpoint before revert to three main chapters
Added bibliography and made minor section changes

--- 2017-12-10 paper.tex
@@
 
-\institute{\email{st152995@stud.uni-stuttgart.de}}
+\institute{\email{aanal.basaula@gmail.com}}
 
@@
 The sheer volume of data generated and the rate at which it is being generated can be termed as Big Data.
-\textbf{Big Data} made it's appearance in 1998 SGI slide by John Mashey and now is of interest to a lot of financial as well as research institutions.[1]
+\textbf{Big Data} made it's appearance in 1998 SGI slide by John Mashey and now is of interest to a lot of financial as well as research institutions.\cite{miningstatus}
 
@@
 A lot of companies utilize pattern recognition on big data to open opportunities for revenue generation whereas research institutions use it to discover patterns simply not observable by the human mind(simply due to the volume of data).
-The question on how to classify any data as big data can be answered by using the 3Vs: Volume, Variety and Velocity [2].
+The question on how to classify any data as big data can be answered by using the 3Vs: Volume, Variety and Velocity \cite{usingapachestorm}.
 Any data with big volume or variety (such as text, video, images, etc) or that gets generated rapidly can be considered as big data.
@@
 
-Apache storm is one of the available open source tools for stream data processing, which is considered to be one of the most easy to use, fast and fault tolerant systems available [3].
+Apache storm is one of the available open source tools for stream data processing, which is considered to be one of the most easy to use, fast and fault tolerant systems available \cite{stormperspective}.
 These three factors have been the major points for the general public to choose \textbf{Apache Storm} above other available tools.
@@
 
-\subsection{Architecture}
+\subsection{Architecture Overview}
 The basic Storm data processing architecture consists of stream of tuples flowing through a defined topology to yield the required analytic result.
@@
 Figure \ref{fig:arch} illustrates the high level architecture where, Nimbus monitors all the available worker nodes and both the nimbus as well as the worker nodes use \textbf{Apache Zookeeper} to store their states.
-The actual processing is done by the \textbf{worker nodes}, and each worker node runs multiple Worker Processes each of them mapped to their own topology [4].
+The actual processing is done by the \textbf{worker nodes}, and each worker node runs multiple Worker Processes each of them mapped to their own topology \cite{stormtwitter}.
 It is also worth while noting that multiple worker processes in a worker node could be executing different parts of the topology.
@@
 Thus, Executors enable intra-topology parallelism whereas Tasks provide intra-spout or intra-bolt parallelism.
-Figure \ref{fig:workerarch} depicts the design of a Worker node.[4]
+Figure \ref{fig:workerarch} depicts the design of a Worker node.\cite{stormtwitter}
 
@@
 Unless specified explicitly, storm assigns single task to a single executor.
-Executors takes tuples from the in queue, examines the task this tuple is destined to, performs the task and then places the result in the out queue [4].
+Executors takes tuples from the in queue, examines the task this tuple is destined to, performs the task and then places the result in the out queue \cite{stormtwitter}.
 
@@
 
-\section{System Design}
-Storm server clusters as discussed above in Section \ref{sec:concepts}, requires at least a Master node and a series of Worker nodes.
-Apart from these, it also requires zookeeper to keep track of the status of all the master as well as the worker nodes.
-Setup of these nodes are trivial and can be configured easily in storm using the configuration files.
-But, the design of a system with regard to the requirements is not such a trivial task because of the different data processing needs.
-The things to consider while designing a Storm Solution is the System architecture i.e.
-Lambda or Lambda-less.
-After selecting one of the above solution, further decisions for the topology design should be considered.
+\section{Stream Processing}
+A stream is a continuous flow of data from one point to the another.
+A Stream Processing system works on these data streams, the work may just be simple transformations, but also could be as complex as predictive analysis.
+A stream processing system depending on the nature of application has different requirements.
+
+Let us consider a Streaming service (may it be video or audio), which performs some transformations on the data stream before serving the client.
+A loss in packets is not a major problem as a little degradation in Quality for a short period of time is acceptable in this scenario.
+As a counter example scenario, consider a service that keeps track of the stream of purchases, it is necessary to not miss any of these important data.
+Apart from this, we also need to consider various other factors for design of a stream processing system, such as resilience, scalability, etc.
+These considerations will be scrutinized in this section.
 
 \subsection{System Architecture}
-Apache storm is a flexible software which is capable of adapting to the requirements.
-It can be used in a lambda architecture or a lambda less architecture.Depending upon the requirements of the system, one can opt for either of the above solutions
+As discussed above, a system depending upon the requirement is allowed to drop a few data sets thus allowing us to utilize the lambda-less architecture, whereas others require complete data sets therefore utilizing the Lambda Architecture.
 
@@
 \paragraph{Apache Storm} can handle data processing without the need of persistence layer and offers its own 
-resiliency mechanisms to make sure data is tuples injected into the system are processed.
-It has two types of processing schemes, \textbf{at least once} where each tuple is processed as the name suggests, at least once, and the other \textbf{at most once}, where the tuples are either processed once or dropped in cases of failure.
-To provide \textbf{at least once} semantics, Storm maintains an acker bolt, which keeps track of the tuple flow in the directed acyclic graph.
+resiliency mechanisms to make sure data is tuples injected into the system are processed. Section \ref{sec:resilience} explains in further detail how storm tries and guarantees a reliable throughput.
 
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=.7\textwidth]{acker.png}
-    \caption{Storm Acker Bolt [4]}
-    \label{fig:acker}
-   \end{center}
-\end{figure}
-
-Figure \ref{fig:acker} shows the Acker bolt integration with the storm topology.
-It keeps track of the tuple flow starting from the point where it has been injected into the topology till it leaves the topology.
-After the tuple has left the topology, an \textbf{ack} is sent to the Spout, at this point, it can decide to remove the Tuple from memory.
+Lambda less architectures are not reliable and do not tend to provide persistence storage as every result is stored in memory and is suitable only for systems which require no storage of calculated information.
+In section \ref{sec:usecase}
 
@@
 
-Data is persisted in the persistence layer so that in cases of high data volume, Batch processing can be done on the input data to provide guaranteed processing and to reduce the computation requirement on the Real Time Data Processing Unit.
+Data is persisted in the persistence layer so that in cases of high data volume, Batch processing can be done on the input data to provide guaranteed processing and to reduce the computation requirement on the Real Time Data Processing Unit.\cite{lambdastorm}
 
@@
 
+\subsection{Storm Resilience}
+\label{sec:resilience}
+
+It has two types of processing schemes, \textbf{at least once} where each tuple is processed as the name suggests, at least once, and the other \textbf{at most once}, where the tuples are either processed once or dropped in cases of failure.
+To provide \textbf{at least once} semantics, Storm maintains an acker bolt, which keeps track of the tuple flow in the directed acyclic graph.
+
 \begin{figure}
   \begin{center}
-    \includegraphics[width=\textwidth]{ranking.png}
-    \caption{Example Ranking Topologies}
-    \label{fig:ranking}
+    \includegraphics[width=.7\textwidth]{acker.png}
+    \caption{Storm Acker Bolt \cite{stormtwitter}}
+    \label{fig:acker}
    \end{center}
@@
 
+\paragraph{Figure} \ref{fig:acker} shows the Acker bolt integration with the storm topology.
+It keeps track of the tuple flow starting from the point where it has been injected into the topology till it leaves the topology.
+After the tuple has left the topology, an \textbf{ack} is sent to the Spout, at this point, it can decide to remove the Tuple from memory.
+
 \section{Practical Use Case}
+\label{sec:usecase}
 Apache Storm is used for stream processing of data, in many sources [2,3] a word frequency count example has been shown which can be considered to be trivial, but in actual practice it can be used to execute complicated algorithms.
@@
 
+ The zookeeper services can be run in a single machine or can be deployed in a cluster as well.
+A single zookeeper instance can also be shared between other services and Apache Storm, but, the major consideration to be taken into account is performance.
+If there are multiple number of systems that utilize the zookeeper, the Zookeeper can run out of number of clients it can service, thus affecting up-time of the systems in general.\cite{stormtwitter} Whereas having a dedicated Zookeeper allows a high performing cluster.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Bibliography}
+\bibliographystyle{splncs03}
+\bibliography{paper}
+Alle Links wurden zuletzt am 22.06.2015 geprüft.
 \end{document}

### 2017-12-30 paper.tex
Made changes to chapter 3

@@
 
-\section{Stream Processing}
+\section{Stream Processing: Design Considerations}
 A stream is a continuous flow of data from one point to the another.
@@
 
-Let us consider a Streaming service (may it be video or audio), which performs some transformations on the data stream before serving the client.
-A loss in packets is not a major problem as a little degradation in Quality for a short period of time is acceptable in this scenario.
-As a counter example scenario, consider a service that keeps track of the stream of purchases, it is necessary to not miss any of these important data.
-Apart from this, we also need to consider various other factors for design of a stream processing system, such as resilience, scalability, etc.
-These considerations will be scrutinized in this section.
+Let us consider the news feed service provided by any social networking site, we know that in any of these websites we can read posts created by our friends or family and it does not matter if we see them at the very instance they post.
+But, there are cases such as a stream of purchases, where we want the data to be visible as soon as available separate design considerations should be made.
+This section focuses on the various considerations that are to be made while designing a stream processing system, specifically using Apache Storm.
 
 \subsection{System Architecture}
-As discussed above, a system depending upon the requirement is allowed to drop a few data sets thus allowing us to utilize the lambda-less architecture, whereas others require complete data sets therefore utilizing the Lambda Architecture.
+As discussed above, a data processing stream has different requirements and according to these various requirements, the System can be modified to suit the needs.
+The major two System architectures currently available are the Lambda Architechture and the Lambda-less (or Kappa) Architecture.
 
 \subsubsection{Lambda-less Architecture}
-\paragraph{Apache Storm} can handle data processing without the need of persistence layer and offers its own 
-resiliency mechanisms to make sure data is tuples injected into the system are processed. Section \ref{sec:resilience} explains in further detail how storm tries and guarantees a reliable throughput.
+\paragraph{Kappa Architecture} or the lambda-less Architecture is a System Architecture concept where there are only two layers, namely Speed Layer and Serving Layer.The data input to the system is processed at real time to retrieve the results at real time.
+The major advantage of this type of system is that there is minimum latency.
+This type of architecture is more suitable in cases where an accurate view of the data is preferable without a lot of delay.
+Cash transactions might be one of these cases where a minimum delay for transaction visualization is suitable added the calculations done on these data sets are properly formulated as well.
 
-Lambda less architectures are not reliable and do not tend to provide persistence storage as every result is stored in memory and is suitable only for systems which require no storage of calculated information.
-In section \ref{sec:usecase}
+This architecture can be considered has an advantage over the lambda architecture (described later) , that a single code base is to be maintained throughout the life of the software.
+But, any apparent change to the code base results in a need for a duplicate cluster to analyze the data again and update the view.
+A working example of this architecture is presented in the section \ref{sec:usecase}.
+One may use this architecture over the lambda architecture when both the algorithms for the batch processing and stream processing are the same and the requirements mentioned above are the major focus.
 
 \subsubsection{Lambda Architecture}
-\paragraph{Lambda Architecture} is a data processing Architecture designed to handle high volume data utilizing both batch and stream processing.
-Apache storm allows users to integrate Persistence models into the system.
-This flexibility allows users to integrate batch processing into the system as well as seen in figure \ref{fig:lambda}.
+\paragraph{Lambda Architecture} is a data processing Architecture designed to handle high volume data utilizing the best of both batch and stream processing.
+Figure \ref{fig:lambda} shows a generic architecture principle for Lambda Architecture.
+The three composed layers are: Speed Layer, Batch Layer and Serving Layer.
+The batch processing layer provides an accurate view of the data but provides results in batches, whereas the speed layer provides current data visualizations which may or may not be accurate depending upon the algorithm.
+The serving layer utilizes both the speed layer as well as the batch layer to fulfill the request to a query.
 
-Data is persisted in the persistence layer so that in cases of high data volume, Batch processing can be done on the input data to provide guaranteed processing and to reduce the computation requirement on the Real Time Data Processing Unit.\cite{lambdastorm}
+Lambda architecture requires maintanence of two separate systems as well as code bases, and thus may result in higher cost.
+But, the added advantage is that the batch processing layer code can be optmized for the given data set to obtain better results.
+An example scenario could be a specialized news feed which requires machine learning algorithms to customize the user experience.
+The Speed layer could be used to deliver recent news, but the batch layer could use machine learning to produce the customized feed.
+An added advantage is that Lambda architecture holds the input data unchanged.\cite{lambdastorm}
 
@@
 \subsection{Topology Design Considerations}
-A storm topology design with the spouts and bolts should be properly considered.
+A storm topology design with the spouts and bolts and affects the overall system thus should be properly considered.
 A spout should inject data into the topology from external sources where as the bolt should be a smallest implementable task.
@@
 \end{figure}
+
 \subsubsection{Resilience:} Apache storm includes APIs which allow users to develop solutions which are resilient in nature.
@@
 Thus, writing a spout which can replay the tuple in case of failures allows us to guarantee the processing of data.
+
 \subsubsection{Interoperability:} The designed topology should be interoperable with different sources of data, and by default should work out of the box by adding simply other Spouts in the topology.
@@
 
+\subsection{Deployment}
+\paragraph{Deploying} a storm topology requires a running zookeeper, nimbus and at least one worker node.
+In this practical implementation we use a single node to handle all the processing tasks.
+The zookeeper as well as the nimbus was instantiated in the local machine, along with a supervisor for the worker node.
+The java code was compiled and packaged as a jar and submitted to nimbus for deployment.
+The topology was left for around 5 hours which can be observed in the screenshots of the reporting UI taken.
+
 \section{Other Tools}
  \label{sec:othertools}
-
+ Stream processing is a topic under constant focus at present and a lot of tools as well as models have been designed and created to support such processing.
+Apart from Apache storm there are other tools such as,
  <Describe shortly that other tools are available as well and list them with short descriptions>
@@
 
- The zookeeper services can be run in a single machine or can be deployed in a cluster as well.
-A single zookeeper instance can also be shared between other services and Apache Storm, but, the major consideration to be taken into account is performance.
-If there are multiple number of systems that utilize the zookeeper, the Zookeeper can run out of number of clients it can service, thus affecting up-time of the systems in general.\cite{stormtwitter} Whereas having a dedicated Zookeeper allows a high performing cluster.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2018-01-10 paper.tex
Fixed Resilience details to concepts section

@@
 Bolts, Spouts.
-We consider a practical implementation of Storm to try and express the theoretical design considerations in real practice.
+We consider a practical implementation of Storm to try and express the theoretical design considerations in practice.
 Finally we list out the other available tools and compare them to Storm with regard to Performance, Resilience and various other parameters.
@@
 
+\subsection{Fault Tolerance and Resilience}
+\label{sec:resilience}
+Apache storm offers fault tolerance and resilience mechanisms with out of the box support and is tolerant against different types of failures.
+Nimbus, being the main master node, is in-charge of all the other worker nodes.
+Under the circumstance that a worker node fails, Nimbus re-spawns a worker node and assigns the tasks to this node.
+If the nimbus itself fails, the nimbus is not re-started automatically, but the worker nodes do not stop processing the stream.
+The nimbus can be later on restarted manually.
+During this phase, if a worker fails, it will not be restarted.
+The supervisor is also an active element within the Storm Architecture and keeps the Nimbus posted about it's current status.
+The \textbf{heartbeat} event is scheduled to run every 15 seconds and it reports to the Nimbus that the worker node is alive.
+\cite{stormtwitter}
+
+Apache Storm supports two types of processing schemes, \textbf{at least once} where each tuple is processed as the name suggests, at least once, and the other \textbf{at most once}, where the tuples are either processed once or dropped in cases of failure.
+To provide \textbf{at least once} semantics, Storm maintains an acker bolt, which keeps track of the tuple flow in the directed acyclic graph.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=.7\textwidth]{acker.png}
+    \caption{Storm Acker Bolt \cite{stormtwitter}}
+    \label{fig:acker}
+   \end{center}
+\end{figure}
+
+\paragraph{Figure} \ref{fig:acker} shows the Acker bolt integration with the storm topology.
+It keeps track of the tuple flow starting from the point where it has been injected into the topology till it leaves the topology.
+After the tuple has left the topology, an \textbf{ack} is sent to the Spout, at this point, it can decide to remove the Tuple from memory.
+
 \section{Stream Processing: Design Considerations}
@@
 As discussed above, a data processing stream has different requirements and according to these various requirements, the System can be modified to suit the needs.
-The major two System architectures currently available are the Lambda Architechture and the Lambda-less (or Kappa) Architecture.
+The major two System architectures currently available are the Lambda Architecture and the Lambda-less (or Kappa) Architecture.
 
@@
 Thus, writing a spout which can replay the tuple in case of failures allows us to guarantee the processing of data.
+For additional details please follow Section \ref{sec:resilience}.
 
@@
 This also amounts to the maintainability of the code.
-
-\subsection{Storm Resilience}
-\label{sec:resilience}
-
-It has two types of processing schemes, \textbf{at least once} where each tuple is processed as the name suggests, at least once, and the other \textbf{at most once}, where the tuples are either processed once or dropped in cases of failure.
-To provide \textbf{at least once} semantics, Storm maintains an acker bolt, which keeps track of the tuple flow in the directed acyclic graph.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=.7\textwidth]{acker.png}
-    \caption{Storm Acker Bolt \cite{stormtwitter}}
-    \label{fig:acker}
-   \end{center}
-\end{figure}
-
-\paragraph{Figure} \ref{fig:acker} shows the Acker bolt integration with the storm topology.
-It keeps track of the tuple flow starting from the point where it has been injected into the topology till it leaves the topology.
-After the tuple has left the topology, an \textbf{ack} is sent to the Spout, at this point, it can decide to remove the Tuple from memory.
 
@@
 
-The only spout uses Twitter Hosebird Client (\url{https://github.com/twitter/hbc}) to connect to Twitter Streaming Server and consume the Tweets generated in real time.
+The only spout available in this example uses Twitter Hosebird Client (\url{https://github.com/twitter/hbc}) to connect to Twitter Streaming Server and consume the Tweets generated in real time.
 Whereas the Bolts were programmed in java and the code can be viewed in \url{https://github.com/hakku276/storm-sentiment-analysis}.
@@
  \label{sec:othertools}
- Stream processing is a topic under constant focus at present and a lot of tools as well as models have been designed and created to support such processing.
-Apart from Apache storm there are other tools such as,
- <Describe shortly that other tools are available as well and list them with short descriptions>
+ Stream processing is a topic under constant focus at present and a lot of tools as well as models have been designed and created to support it. Apart from Apache storm there are other tools such as:
+
+ \begin{enumerate}
+ \item \textbf{Spark Streaming} is an Open Source based project for stream processing.
+It relies on cluster computing systems and operates on batches.
+The batch window in Spark Streaming is comparatively small thus allowing stream processing.It also provides flexibility to work in different programming languages such as Java, Scala and Python.
+ \item \textbf{Apache Samza} 
+ \item \textbf{Yahoo S4} is also a Real time stream processing tool inspired by the MapReduce Model.
+ \end{enumerate}
 
@@
 
+ \subsubsection{Benchmarks}
+
  \section{Discussion}
@@
 \bibliography{paper}
-Alle Links wurden zuletzt am 22.06.2015 geprüft.
+All Links were verified at 10.01.2018
 \end{document}

### 2018-01-12 paper.tex
Starting with Last section

@@
 The zookeeper as well as the nimbus was instantiated in the local machine, along with a supervisor for the worker node.
-The java code was compiled and packaged as a jar and submitted to nimbus for deployment.
-The topology was left for around 5 hours which can be observed in the screenshots of the reporting UI taken.
+The Storm User interface presents all these information in a comprehensible manner within a web browser, which can be observed in figure \ref{fig:stormui} The java code was compiled and packaged as a jar and submitted to nimbus for deployment (refer GitHub readme for further details).
+It was then deployed in storm, which can be observed in the Screen shot for the user interface in figure \ref{fig:stormui}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{ui-mainpage.png}
+    \caption{Main Page for Storm UI, displays all the cluster related details}
+    \label{fig:stormui}
+   \end{center}
+\end{figure}
 
@@
  \label{sec:othertools}
- Stream processing is a topic under constant focus at present and a lot of tools as well as models have been designed and created to support it. Apart from Apache storm there are other tools such as:
+Stream processing is a topic under constant focus at present and a lot of tools as well as models have been designed and created to support it.
+This section makes a comparison of the available widely accepted tool today.
+The available tools are mentioned below in brief such as:
 
@@
 The batch window in Spark Streaming is comparatively small thus allowing stream processing.It also provides flexibility to work in different programming languages such as Java, Scala and Python.
- \item \textbf{Apache Samza} 
- \item \textbf{Yahoo S4} is also a Real time stream processing tool inspired by the MapReduce Model.
+ \item \textbf{Apache Samza} is an Open Source project for near realtime distributed stream processing framework.
+It uses Apache Kafka for messaging and Apache Hadoop YARN for fault tolerance, processor isolation, security and resource management.\footnote{http://samza.apache.org/}
+ \item \textbf{Flink Streaming} is also a Real time stream processing tool for distributed systems and builds batch processing on top of the Streaming engine.
  \end{enumerate}

### 2018-01-17 paper.tex
Added Comparision Section

@@
 
- \subsection{Comparision between Storm, Spark and Samza}
+ \subsection{Comparison between Storm, Spark and Samza}
+This sub section compares the three most widely used tools for Stream Processing, i.e. Storm, Spark and Samza.
 
- <Gather and show the differences between Storm, Spark and Samza in general>
-
- \subsubsection{Coding}
- <Compare how the coding is in all these platforms>
+ \subsubsection{Development and Testing}
+All three tools provide their own API for development and testing.
+A comparision among these tools can be made on a basis of the programming languages supported, Apache Samza currently only supports JVM languages, whereas Spark Streaming supports Java, Scala and Python.
+Apache Storm on the other hand supports JVM as well as Non JVM languages which increases development speed, as the developers can choose their language according to their comfort.
 
  \subsubsection{Tool setup}
- <Compare how easy it is to setup the tools>
+Setting up of the tool in Apache Storm required minimum effort, achieved by downloading the zookeeper and storm binaries and unpacking them onto the computer.
+A minor configuration changes required attention, such as the location where zookeeper and storm would store data in.
+Apart from these details, it worked perfectly out of the box.
+Along the same line, Spark Streaming follows a similar approach for tool setup and can be installed by simply downloading the tarball and unpacking it.
+Apache Samza on the other hand maintains a repository \url(https://github.com/apache/samza-hello-samza) which contains the scripts which installs and starts the requirement dependencies in the host.
 
  \subsubsection{Deployment}
- <Compare how easy it is to deploy the code onto the tool>
+Deployment of an artifact on Apache storm is relatively easy.
+For the java application, one compiles the sources with all the required dependencies into a jar file.
+The jar file can then be easily deployed into storm via a simple command.
 
- \subsubsection{Resilience}
- <Compare the resilience model of the tools>
+ \subsubsection{Fault Tolerance}
+As stream processing systems run for extended periods, it is quite necessary that these should be able to handle various kinds of faults and still keep processing data as they enter the system.
+The fault tolerance model for Storm has been extensively described in section \ref{sec:resilience}.
+Similarly, Spark Streaming and Apache Samza also use various techniques to resolve issues while deployed.
+Spark Streaming uses write ahead logs(journals) to keep track of all the data operation, and during failure of a worker, the worker is restarted and a state restore is done by reading the logs.
+This ensures that the data is processed as required.
+In comparison to Apache Storm, Storm continues to process incoming tuples even when the master node fails, but in Spark, the worker nodes stop working when the master node fails until a new leader starts.
+This is an added bonus to Storm, since a failed master manual restart would reinstate the state of the system without many changes.
 
  \subsubsection{Benchmarks}
+There have not been a comparative benchmark between Apache Storm, Spark Streaming and Apache Samza, but various sources have different comparisions between tools.
+\cite{benchmark} considers Apache Storm, Spark Streaming and Apache Flink.
+The benchmark design consisted of counting the number of events taking place in the stream.
+The design was setup to read data from a Kafka Event Queue, the data being processed and result updated into Redis.
+The result included the window start time and the window end time, thus giving  an idea about the total latency of processing.
+According to this benchmark, they deduced that Apache Storm and Flink had low latency, whereas Spark Streaming had higher maximum throughput depending upon the batch duration setting.
 
  \section{Discussion}
-<What I felt about the tool, ease of use and other general aspects>
 

### 2018-01-22 paper.tex
Prepare for review with supervisor

@@
 
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=.7\textwidth]{topo.png}
-    \caption{Generic Storm Topology}
-    \label{fig:topo}
-   \end{center}
-\end{figure}
-
 \subsubsection{Spout} 
@@
 It receives raw tuples emitted by spouts or processed tuples sent by another bolt and sends it down the pipe after it has completed its processing.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.85\textwidth, height= 0.23\textheight]{topo.png}
+    \caption{Generic Storm Topology}
+    \label{fig:topo}
+   \end{center}
+\end{figure}
 
@@
 \cite{stormtwitter}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth, height=0.25\textheight]{throughput.png}
+    \caption{Throughput in time with respect to changing worker node availability \cite{stormtwitter}}
+    \label{fig:throughput}
+   \end{center}
+\end{figure}
+
+As we can see in figure \ref{fig:throughput}, an experiment conducted in \cite{stormtwitter} to check the resilience and performance of storm under loss of worker nodes.
+The worker nodes were turned off one by one in fixed durations, and the throughput of the system was measured at every instance.
+The durations and number of machines are shown below in table \ref{tab:experiment}.
+As seen, 3 machines were killed every 3 minutes, but the number of workers were limited to 50 at every instance, thus increasing the number of workers per machine.
+The graph \ref{fig:throughput} suggests that every 15 minutes, when the worker nodes were killed, the throughput was reduced significantly while storm tries to balance the topology again.
+The balancing does not require extensible amount of time and results into a new maximum throughput.
+As we can observe in the graph, the throughput decreses every time a node is removed, which is an expected phenomenon, due to the decrease in the computational capacity.
+
+\begin{table}
+\caption{Experimental setup during different time frames}
+\label{tab:experiment}
+\begin{center}
+\begin{tabular}{r@{\quad}cl}
+\hline
+\multicolumn{1}{l}{\rule{0pt}{12pt} 
+	Time}&\multicolumn{1}{l}{no of machines}\\[2pt]
+\hline\rule{0pt}{12pt}
+0 minutes  & 16  \\
++15 minutes & 13 \\
++30 minutes & 10 \\
++45 minutes & 7 \\
++60 mintues & 4 \\[2pt]
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 Apache Storm on the other hand supports JVM as well as Non JVM languages which increases development speed, as the developers can choose their language according to their comfort.
+Apart from the fact that choosing a language for development, users can utilize existing code, thus allowing high portability.
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=.75 \textwidth]{comparisiongraph.png}
+\caption{Performance Comparision between Apache Storm, Apache Flink and Spark Streaming \cite{benchmark}}
+\label{fig:comparisiongraph}
+\end{center}
+\end{figure}
+
  \subsubsection{Benchmarks}
@@
 According to this benchmark, they deduced that Apache Storm and Flink had low latency, whereas Spark Streaming had higher maximum throughput depending upon the batch duration setting.
+Figure \ref{fig:comparisiongraph} shows the comparative analysis among Apache Flink, Spark Streaming and Apache Storm.
+The batch processing scheme of Spark Streaming allows it to process events with higher throughput but with a trade off that the processing is delayed with higher latency.
+In the figure, we can observe that storm has comparatively less delay in processing, but according to \cite{benchmark}, storm with acking beyond 135,000 events per second, could not keep up with the throughput.
 
  \section{Discussion}
+Apache Storm acts as a real time stream processing tool which is capable of computing results with high reliability and minimum latency.
+The tool is comparatively easy to setup and configure, with out of the box configurations which is production ready.
+The storm API is relatively simple and does not have a complicated method calls, the programming is based on event driven programming and since storm supports read and writes based on standard IO, utilizing previously written code could be relatively easier than in the other available tools.
+Data is currently being generated and consumed at an unprecedented rate, thus stream processing technologies is a requirement and is gaining popularity and attention.
+With all the various tools available, one can decide on which tool best suites ones needs.
+According to the benchmarks, if minimum latency is required, Apache Storm or Apache Flink or any system with Continuous data flow model should be preferred.
+Whereas if high throughput is required with a non strict latency criteria, Spark streaming could prove to be a better option.
 

### 2018-01-25 paper.tex
Minor changes to report as suggested

@@
 The question on how to classify any data as big data can be answered by using the 3Vs: Volume, Variety and Velocity \cite{usingapachestorm}.
-Any data with big volume or variety (such as text, video, images, etc) or that gets generated rapidly can be considered as big data.
+Any data with big volume or variety (such as text, video, images, etc.) or that gets generated rapidly can be considered as big data.
 In this paper regarding \textbf{Apache Storm} we consider the third V of big data to be the influencing factor.
@@
 In all of these scenarios, the data is continuously being generated.
-Traditionally analyzing of these types of data requires storage of these data onto persistence and later batch processing it to discern patterns as well as other useful information.
-But, with the help of \textbf{Apache Storm} we are able to process these data as they are generated with minimum amount of latency.
+Traditionally analyzing these types of data requires storage of these data onto persistence and later batch processing it to discern patterns as well as other useful information.
+But, with the help of \textbf{Apache Storm} we are able to process these data as they are generated with minimum latency.
 
@@
 
-\paragraph{The Master node} also called the \textbf{Nimbus} is also responsible for distribution and coordination of the execution of topologies.
+\paragraph{The Master node} also called the \textbf{Nimbus} is also responsible for the distribution and coordination of the execution of topologies.
 The Nimbus is the first point where the user submits the defined topology, which is then distributed among the worker nodes.
@@
 The actual processing is done by the \textbf{worker nodes}, and each worker node runs multiple Worker Processes each of them mapped to their own topology \cite{stormtwitter}.
-It is also worth while noting that multiple worker processes in a worker node could be executing different parts of the topology.
+It is also worth noting that multiple worker processes in a worker node could be executing different parts of the topology.
 
 \subsubsection{Worker Node}
-\paragraph{Worker Node} is responsible for all the heavy lifting required for processing of the topology.
+\paragraph{Worker Node} is responsible for all the heavy lifting required for processing the topology.
 One worker node may run a multiple number of worker processes.
@@
 \paragraph{Supervisor} is running in every Worker Node and communicates with the Nimbus to receive assignments from it.
-Apart from taking assignments from the Nimbus it is also responsible for monitoring the health of the of the worker processes and respawns the worker processes in case it dies.
+Apart from taking assignments from the Nimbus it is also responsible for monitoring the health of the worker processes and respawns the worker processes in case it dies.
 
@@
 Unless specified explicitly, storm assigns single task to a single executor.
-Executors takes tuples from the in queue, examines the task this tuple is destined to, performs the task and then places the result in the out queue \cite{stormtwitter}.
+Executors take tuples from the in queue, examines the task this tuple is destined to, performs the task and then places the result in the out queue \cite{stormtwitter}.
 
 \subsection{Topology}
-\paragraph{Topology} is a directed graph where the vertices represent computation and the edges represents the data flow.
+\paragraph{Topology} is a directed graph where the vertices represent computation and the edges represent the data flow.
 Another important thing about topologies in Storm is that they are allowed to have cycles.
@@
   \begin{center}
-    \includegraphics[width=0.85\textwidth, height= 0.23\textheight]{topo.png}
+    \includegraphics[width=0.85\textwidth]{topo.png}
     \caption{Generic Storm Topology}
@@
 The balancing does not require extensible amount of time and results into a new maximum throughput.
-As we can observe in the graph, the throughput decreses every time a node is removed, which is an expected phenomenon, due to the decrease in the computational capacity.
+As we can observe in the graph, the throughput decreases every time a node is removed, which is an expected phenomenon, due to the decrease in the computational capacity.
 
@@
 
-Apache Storm supports two types of processing schemes, \textbf{at least once} where each tuple is processed as the name suggests, at least once, and the other \textbf{at most once}, where the tuples are either processed once or dropped in cases of failure.
+Apache Storm supports two types of processing schemes, \textbf{at least once} where each tuple is processed as the name suggests, and the other \textbf{at most once}, where the tuples are either processed once or dropped in cases of failure.
 To provide \textbf{at least once} semantics, Storm maintains an acker bolt, which keeps track of the tuple flow in the directed acyclic graph.
@@
 	\section{Stream Processing: Design Considerations}
-A stream is a continuous flow of data from one point to the another.
-A Stream Processing system works on these data streams, the work may just be simple transformations, but also could be as complex as predictive analysis.
+A stream is a continuous flow of data from one point to the other.
+A Stream Processing system works on these data streams, the work may just be simple transformations, but could also be as complex as predictive analysis.
 A stream processing system depending on the nature of application has different requirements.
@@
 \subsection{System Architecture}
-As discussed above, a data processing stream has different requirements and according to these various requirements, the System can be modified to suit the needs.
+As discussed above, a data processing stream has different requirements and according to these various requirements, the System can be modified to suit these needs.
 The major two System architectures currently available are the Lambda Architecture and the Lambda-less (or Kappa) Architecture.
@@
 A storm topology design with the spouts and bolts and affects the overall system thus should be properly considered.
-A spout should inject data into the topology from external sources where as the bolt should be a smallest implementable task.
+A spout should inject data into the topology from external sources whereas the bolt should be a smallest implementable task.
 A storm topology should also be considered with various considerations in mind:
@@
 Since we should have only one instance of Aggregator, to be able to get the aggregated result, in cases of increased data load, the number of rolling count bolts can be increased, but the Aggregation Bolt cannot be increases thus adding increased load at this node.
-To relax the demand on this node, we can introduce another bolt, the Intermediate Ranking Bolt, which is responsible for ranking the words at a certain predefined intervals and passes the intermediate ranks to the aggregator.
+To relax the demand on this node, we can introduce another bolt, the Intermediate Ranking Bolt, which is responsible for ranking the words at certain predefined intervals and passes the intermediate ranks to the aggregator.
 This reduces load on the Aggregator since we can increase the intermediate ranking bolts to cope with the increased load.
@@
 \subsubsection{Resilience:} Apache storm includes APIs which allow users to develop solutions which are resilient in nature.
-Using the acker bolt as show in figure \ref{fig:acker}, users can track the progress of the tuple injected into the system.
+Using the acker bolt as shown in figure \ref{fig:acker}, users can track the progress of the tuple injected into the system.
 Thus, writing a spout which can replay the tuple in case of failures allows us to guarantee the processing of data.
@@
 The company can also analyze the feedback people place on a specific product, to change and adjust the pricing schemes in real time to the demand.
-At present there are many social media networks available and people tend to share their views on these networks,  \textbf{Twitter} is also one of these social networks where a number of posts are available publicly and can be analyzed to obtain a generalized view about a specific product or the company in general.
+At present there are many social media networks available and people tend to share their views on these networks,  \textbf{Twitter}, for instance, is also one of these social networks where a number of posts are available publicly and can be analyzed to obtain a generalized view about a specific product or the company in general.
 
@@
 \paragraph{This} is not the only possible solution, there are multiple other tools such as \textbf{Spark Streaming}, \textbf{Samza}, etc.
-but this paper considers Apache Storm as it is in the spotlight of this paper, for further details regarding other available tools and their comparisons with Apache Storm, please consider the section \ref{sec:othertools} for further details.
+but this paper concentrates Apache Storm as it is in the spotlight of this paper, for further details regarding other available tools and their comparisons with Apache Storm, please consider the section \ref{sec:othertools} for further details.
 
@@
 In the design phase, the storm topology was thoroughly planned, specifically the spouts, bolts and how they are interconnected.
-In the development phase, these spouts, bolts and other required aspects of the software was programmed and lastly was deployed on a locally hosted storm server.
+In the development phase, these spouts, bolts and other required aspects of the software were programmed and lastly deployed on a locally hosted storm server.
 
@@
 The ParseTweetBolt processes these tweets and then emits whether the tweet was positive or negative under the simplified scenario.
-These separate positive or negative remark on a tweet does not provide useful information unless it is aggregated, thus a ResultAggregationBolt is responsible for aggregating all the results of the Parser bolt and providing the result to the target in real time.
+This separate positive or negative remark on a tweet does not provide useful information unless it is aggregated, thus a ResultAggregationBolt is responsible for aggregating all the results of the Parser bolt and providing the result to the target in real time.
 
@@
 The zookeeper as well as the nimbus was instantiated in the local machine, along with a supervisor for the worker node.
-The Storm User interface presents all these information in a comprehensible manner within a web browser, which can be observed in figure \ref{fig:stormui} The java code was compiled and packaged as a jar and submitted to nimbus for deployment (refer GitHub readme for further details).
+The Storm User interface presents all these information in a comprehensible manner within a web browser, which can be observed in figure \ref{fig:stormui}, the java code was compiled and packaged as a jar and submitted to nimbus for deployment (refer GitHub readme for further details).
 It was then deployed in storm, which can be observed in the Screen shot for the user interface in figure \ref{fig:stormui}.
@@
 Stream processing is a topic under constant focus at present and a lot of tools as well as models have been designed and created to support it.
-This section makes a comparison of the available widely accepted tool today.
+This section makes a comparison of the available widely accepted tools today.
 The available tools are mentioned below in brief such as:
@@
 It relies on cluster computing systems and operates on batches.
-The batch window in Spark Streaming is comparatively small thus allowing stream processing.It also provides flexibility to work in different programming languages such as Java, Scala and Python.
+The batch window in Spark Streaming is comparatively small thus allowing stream processing.
+It also provides flexibility to work in different programming languages such as Java, Scala and Python.
  \item \textbf{Apache Samza} is an Open Source project for near realtime distributed stream processing framework.
@@
 Apache Storm on the other hand supports JVM as well as Non JVM languages which increases development speed, as the developers can choose their language according to their comfort.
-Apart from the fact that choosing a language for development, users can utilize existing code, thus allowing high portability.
+Apart from being able to choose a language for development, users can utilize existing code, thus allowing high portability.
 
  \subsubsection{Tool setup}
-Setting up of the tool in Apache Storm required minimum effort, achieved by downloading the zookeeper and storm binaries and unpacking them onto the computer.
+Setting up of the tool in Apache Storm requires minimum effort, achieved by downloading the zookeeper and storm binaries and unpacking them onto the computer.
 A minor configuration changes required attention, such as the location where zookeeper and storm would store data in.
-Apart from these details, it worked perfectly out of the box.
+Apart from these details, it works perfectly out of the box.
 Along the same line, Spark Streaming follows a similar approach for tool setup and can be installed by simply downloading the tarball and unpacking it.
@@
  \subsubsection{Fault Tolerance}
-As stream processing systems run for extended periods, it is quite necessary that these should be able to handle various kinds of faults and still keep processing data as they enter the system.
+As stream processing systems run for extended periods, it is quite necessary that these are able to handle various kinds of faults and still keep processing data as they enter the system.
 The fault tolerance model for Storm has been extensively described in section \ref{sec:resilience}.
@@
  \subsubsection{Benchmarks}
-There have not been a comparative benchmark between Apache Storm, Spark Streaming and Apache Samza, but various sources have different comparisions between tools.
+There has not been a comparative benchmark between Apache Storm, Spark Streaming and Apache Samza, but various sources have different comparisions between tools.
 \cite{benchmark} considers Apache Storm, Spark Streaming and Apache Flink.

### 2018-01-31 paper.tex
Begin changes as per suggestion

@@
 The sheer volume of data generated and the rate at which it is being generated can be termed as Big Data.
-\textbf{Big Data} made it's appearance in 1998 SGI slide by John Mashey and now is of interest to a lot of financial as well as research institutions.\cite{miningstatus}
+The word \textbf{Big Data} made it's appearance in 1998 SGI slide deck made by John Mashey and now is of interest to a lot of financial as well as research institutions.\cite{miningstatus}
 
@@
 
+\section{Stream Processing}
+A stream is a continuous flow of data from one point to the other.
+Stream Processing System works on these data streams, the task of these processing scheme may range from simple transformations to something as complex as event processing.
+In this present digitized world, people tend to post most of their thoughts and voices about topics on the web.
+These may be in the form of a long text, generally a webpage called blogs, or as a short post on a social networking site.
+An example of a social networking site is Twitter.
+In twitter, people tend to post short views about various things and at any moment there are a number of people posting onto twitter.
+
+A company which is interested in analyzing the current outlook of the organization can decide to analyze these posts on twitter.
+The posts created by twitter users could now then be considered as Data Stream for the company.
+As seen in \ref{fig:example} various users generate posts on twitter, which is stored on the twitter servers.
+These tweets are used by the company to analyze the sentiment of the users and the Reporting UI represents the results to the respective interested party.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{example.png}
+    \caption{An Example Stream Processing Scenario}
+    \label{fig:example}
+   \end{center}
+\end{figure}
+
+\subsection{Stream Processing Models}
+There are two types of stream processing models, namely continuous flow model and discretized stream model. \cite{streaming}
+
+\subsubsection{Continuous Stream Processing}
+
+\subsubsection{Discretized Stream Processing}
+
 \section{Apache Storm: Concepts}
@@
 
-Apache storm is one of the available open source tools for stream data processing, which is considered to be one of the most easy to use, fast and fault tolerant systems available \cite{stormperspective}.
+Apache storm is one of the available open source tools for stream data processing, which is considered to be one of the simplest, fastest and fault tolerant systems available \cite{stormperspective}.
 These three factors have been the major points for the general public to choose \textbf{Apache Storm} above other available tools.
-It was originally created by Nathan Marz at BackType and till now, at present more than 60 companies use or experiment with this tool.
+It was originally created by Nathan Marz at BackType and till now, at present more than 60 companies \cite{stormtwitter} use or experiment with this tool.
 
 \subsection{Architecture Overview}
-The basic Storm data processing architecture consists of stream of tuples flowing through a defined topology to yield the required analytic result.
-Processing of this stream data is performed in Nodes called the worker nodes, which are responsible for performing the required calculations.
-There is also another node named the Master Node, which is responsible for the complete Storm infrastructure,  by keeping track of all the worker nodes and the total topology itself.
+The physical high level architecture of Storm as seen in figure \ref{fig:physicalarc} consists of a single master node and multiple worker nodes.
+The master node(see later) is the leader of the cluster and is responsible in short for the management of the different parts of the topology.
+There can only be one master node in a cluster.
+The worker node on the other hand is responsible for the actual execution of tasks supplied by the user.
+A user task is defined in storm as a Topology which actually is a directed graph.
+The vertices of the graph represent the data processing being done and the edges the flow of data.
+A data stream passed through the topology yields the required processing result as defined by the user.
 
@@
     \includegraphics[width=.7\textwidth]{arch.png}
-    \caption{High Level Storm Architecture}
-    \label{fig:arch}
+    \caption{High Level Storm Physical Architecture}
+    \label{fig:physicalarc}
    \end{center}
@@
 
-\subsubsection{Nimbus}
+The logical or the runtime architecture of storm is defined within a topology.
+As seen in figure \ref{fig:logicalarch} shows an example of a topology and consists of vertices and edges, the vertices represent point where either data is generated or some processing to the existing data is done.
+The edges represent the flow of data, and in a topology, the data is represented in the form of tuples.
+A vertex can in storm be either a spout or a bolt.
+A spout is the source of tuples and it injects tuples into the topology, wheras a bolt is a vertice within a topology where the injected tuples are processed.
 
-\paragraph{The Master node} also called the \textbf{Nimbus} is also responsible for the distribution and coordination of the execution of topologies.
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=.7\textwidth]{arch.png}
+    \caption{High Level Storm run-time or Logical Architecture}
+    \label{fig:logicalarch}
+   \end{center}
+\end{figure}
+
+The logical and physical architectures meet while being deployed in a cluster, the worker nodes are responsible for executing the topology.
+Every worker spawns a number of processes which run topologies.
+See later for detailed explanation on how topologies can be divided and executed within a storm cluster.
+
+\subsubsection{Storm Components}
+\paragraph{Master node} also called the \textbf{Nimbus} is responsible for the distribution and coordination of the execution of topologies.
 The Nimbus is the first point where the user submits the defined topology, which is then distributed among the worker nodes.
-Figure \ref{fig:arch} illustrates the high level architecture where, Nimbus monitors all the available worker nodes and both the nimbus as well as the worker nodes use \textbf{Apache Zookeeper} to store their states.
-The actual processing is done by the \textbf{worker nodes}, and each worker node runs multiple Worker Processes each of them mapped to their own topology \cite{stormtwitter}.
-It is also worth noting that multiple worker processes in a worker node could be executing different parts of the topology.
+The user submits the code as a jar file to the nimbus, which in turn uses a combination of local disk and the zookeeper to store the topology defined in the jar file.\cite{stormtwitter} The topology thrift objects are stored in the zookeeper, whereas the code itself is stored in the local disk of nimbus.
+\paragraph{Worker Node} is responsible for all the heavy lifting required for processing the incoming data stream.
+A worker node maintains a \textbf{Supervisor} process and multiple worker processes.
+The number of  worker processes depends upon the storm configurations.
+A worker process once spawned, starts working for a single topology alone and not for a multiple number of topologies.
+It creates threads called \textbf{Executors} which take up a tasks to be performed.
+A task for an executor is the vertex of the Directed graph, ie.
+either a bolt or a spout.
+This can be observed in the \ref{fig:workerarch}, where a worker node has multiple worker processes, which in turn has multiple executors running different tasks.
+Even though storm can run multiple topologies, different processes run separate topologies.
 
-\subsubsection{Worker Node}
-\paragraph{Worker Node} is responsible for all the heavy lifting required for processing the topology.
-One worker node may run a multiple number of worker processes.
-Each of these worker processes may belong to the same topology or different topologies.
-A worker node maintains a \textbf{Supervisor} and a multiple number of Worker Processes which in turn maintains a multiple number of executors.
-Thus, Executors enable intra-topology parallelism whereas Tasks provide intra-spout or intra-bolt parallelism.
-Figure \ref{fig:workerarch} depicts the design of a Worker node.\cite{stormtwitter}
+The introduction of tasks enables a parallel execution of a number of same bolt or spout node of a topology, whereas the introduction of executors enables us to run different parts of the topology namely bolts and spouts in parallel.\cite{stormtwitter}.
+It is worth noting that these different parts of the topology, namely bolts and spouts can be executing within a single worker process in different executors, or can also be in different executors across different worker nodes.
 
@@
 
-\paragraph{Supervisor} is running in every Worker Node and communicates with the Nimbus to receive assignments from it.
-Apart from taking assignments from the Nimbus it is also responsible for monitoring the health of the worker processes and respawns the worker processes in case it dies.
+\paragraph{Supervisor} is part of a Worker Node and communicates with Nimbus and accepts incoming jobs.
+It monitors the health of the worker processes and respawns every process in case of failure.
+The supervisor performs three major periodic tasks, namely heartbeat, synchronize supervisor and sunchronize process.
+The \textbf{heartbeat} event is scheduled every 15 seconds by the supervisor and in this event it reports to the nimbus that the worker node is alive and provides information about the available worker processes.\cite{stormtwitter} With this provided information, nimbus decides whether it can accept new jobs or not and depending upon the requirements, schedules jobs to be performed by the node.
+The \textbf{synchronize supervisor} event runs every 10 seconds and is reponsible for synchronization of the supervisor.
+It monitors the change is existing assignments and if the change includes an addition of topology, then downloads the required jar files and its dependencies and immediately schedules a synchronize worker process event.
 
-\paragraph{Worker Process} is the Entity that does the processing and is mapped to a certain topology.
-Each worker node spawns a JVM and runs one or multiple executors.
-These executors may be running any of the bolt or spout within the topology, thus providing an added level of parallelism.
+The \textbf{synchronize worker process} event runs every 3 seconds and is responsible for managing worker processes.
+It reads worker heartbeats and classifies those workers with one of the labels, \textit{valid, timed out, not started or disallowed}.
 
-\paragraph{Executors} are made up of one or multiple Tasks.
-Unless specified explicitly, storm assigns single task to a single executor.
-Executors take tuples from the in queue, examines the task this tuple is destined to, performs the task and then places the result in the out queue \cite{stormtwitter}.
-
-\subsection{Topology}
+\subsubsection{Topology}
 \paragraph{Topology} is a directed graph where the vertices represent computation and the edges represent the data flow.
@@
 
-\subsubsection{Spout} 
 \paragraph{Spouts} are the entry point for any data in the Storm Topology.
@@
 
-\subsubsection{Bolt} 
 \paragraph{Bolts} are the processing entities of a Topology.
@@
 
-\subsection{Fault Tolerance and Resilience}
+\subsection{Storm Features}
+\subsubsection{Fault Tolerance and Resilience}
 \label{sec:resilience}
@@
 \paragraph{Solution} to the scenario can be implemented using \textbf{Apache Storm} to process the tweets in real time.
-Figure \ref{fig:solution} shows the simplified system overview, consisting of Twitter Streaming Servers as the source of information.
+Figure \ref{fig:example} shows the simplified system overview, consisting of Twitter Streaming Servers as the source of information.
 This information is processed in the Storm Clusters from within the Company Infrastructure and the analyzed result is reported to the User Interface, which can be accessed by the appropriate user.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=.7\textwidth]{solution.png}
-    \caption{Generic Streaming Solution implemented in this use case}
-    \label{fig:solution}
-   \end{center}
-\end{figure}
 

### 2018-02-18 paper.tex
Fixes made according to final request

@@
 
-\section{Stream Processing}
+\section{Apache Storm: Concepts}
+\label{sec:concepts}
 A stream is a continuous flow of data from one point to the other.
@@
 
-\subsection{Stream Processing Models}
-There are two types of stream processing models, namely continuous flow model and discretized stream model. \cite{streaming}
-
-\subsubsection{Continuous Stream Processing}
-
-\subsubsection{Discretized Stream Processing}
-
-\section{Apache Storm: Concepts}
-\label{sec:concepts}
-
 Apache storm is one of the available open source tools for stream data processing, which is considered to be one of the simplest, fastest and fault tolerant systems available \cite{stormperspective}.
@@
 There can only be one master node in a cluster.
-The worker node on the other hand is responsible for the actual execution of tasks supplied by the user.
-A user task is defined in storm as a Topology which actually is a directed graph.
-The vertices of the graph represent the data processing being done and the edges the flow of data.
-A data stream passed through the topology yields the required processing result as defined by the user.
+The worker node on the other hand is responsible for the actual execution of user logic.
+Storm requires at least one worker node to execute the user logic and apart from it a Zookeeper to keep track of the state of the various nodees in the cluster.
 
@@
     \includegraphics[width=.7\textwidth]{arch.png}
-    \caption{High Level Storm Physical Architecture}
+    \caption{Storm Physical Architecture}
     \label{fig:physicalarc}
@@
 
-The logical or the runtime architecture of storm is defined within a topology.
-As seen in figure \ref{fig:logicalarch} shows an example of a topology and consists of vertices and edges, the vertices represent point where either data is generated or some processing to the existing data is done.
-The edges represent the flow of data, and in a topology, the data is represented in the form of tuples.
-A vertex can in storm be either a spout or a bolt.
-A spout is the source of tuples and it injects tuples into the topology, wheras a bolt is a vertice within a topology where the injected tuples are processed.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=.7\textwidth]{arch.png}
-    \caption{High Level Storm run-time or Logical Architecture}
-    \label{fig:logicalarch}
-   \end{center}
-\end{figure}
-
-The logical and physical architectures meet while being deployed in a cluster, the worker nodes are responsible for executing the topology.
-Every worker spawns a number of processes which run topologies.
-See later for detailed explanation on how topologies can be divided and executed within a storm cluster.
+The logical or the runtime architecture of storm is defined in storm via the abstraction of a topology.
+Topoloyg is a directed graph, with data flows and transformations to the flowing data.
+As seen in figure \ref{fig:topo} shows an example of a topology and consists of vertices and edges, the edges represent the flow of data whereas the vertices represent point where either data is generated or some processing to the existing data flowing throw it.
 
@@
 The Nimbus is the first point where the user submits the defined topology, which is then distributed among the worker nodes.
-The user submits the code as a jar file to the nimbus, which in turn uses a combination of local disk and the zookeeper to store the topology defined in the jar file.\cite{stormtwitter} The topology thrift objects are stored in the zookeeper, whereas the code itself is stored in the local disk of nimbus.
-\paragraph{Worker Node} is responsible for all the heavy lifting required for processing the incoming data stream.
+The user submits the code as a jar file to the nimbus, nimbus then uses a combination of local disk and the zookeeper to store the topology defined in the jar file.\cite{stormtwitter} The topology thrift objects are stored in the zookeeper, whereas the code itself is stored in the local disk of nimbus.
+It then assigns the task to the worker nodes to execute selected parts of the topology.
+\paragraph{Worker Node} is responsible for all the execution required for processing the incoming data stream.
 A worker node maintains a \textbf{Supervisor} process and multiple worker processes.
-The number of  worker processes depends upon the storm configurations.
+The number of  worker processes depends upon the storm configurations and the number of woker nodes available.
 A worker process once spawned, starts working for a single topology alone and not for a multiple number of topologies.
 It creates threads called \textbf{Executors} which take up a tasks to be performed.
-A task for an executor is the vertex of the Directed graph, ie.
-either a bolt or a spout.
+A task for an executor is the vertex of the Directed graph.
 This can be observed in the \ref{fig:workerarch}, where a worker node has multiple worker processes, which in turn has multiple executors running different tasks.
-Even though storm can run multiple topologies, different processes run separate topologies.
-
-The introduction of tasks enables a parallel execution of a number of same bolt or spout node of a topology, whereas the introduction of executors enables us to run different parts of the topology namely bolts and spouts in parallel.\cite{stormtwitter}.
-It is worth noting that these different parts of the topology, namely bolts and spouts can be executing within a single worker process in different executors, or can also be in different executors across different worker nodes.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=.7\textwidth]{worker.png}
-    \caption{Worker Node Architecture}
-    \label{fig:workerarch}
-   \end{center}
-\end{figure}
+Even though storm can run multiple topologies, different processes may run separate topologies.
 
@@
 
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=.7\textwidth]{worker.png}
+    \caption{Worker Node Architecture}
+    \label{fig:workerarch}
+   \end{center}
+\end{figure}
+
 \subsubsection{Topology}
-\paragraph{Topology} is a directed graph where the vertices represent computation and the edges represent the data flow.
-Another important thing about topologies in Storm is that they are allowed to have cycles.
-Figure \ref{fig:topo} shows a basic Storm Topology, which consists of Spouts and Bolts.
-Spouts are the Sources of tuples in a Topology whereas Bolts are the units which process these generated tuples.
-The user first submits the topology to the Apache Storm Nimbus, which consults with each of the available Supervisors in the worker nodes and assigns each of them with tasks related to the topology.
+\paragraph{Topology} is a directed graph where as mentioned before the vertices represent computation and the edges represent the data flow.
+It is wise to note that topologies in Storm are allowed to have cycles.
+The data in storm is abstracted using a Data type called the Tuple.
+A tuple is a list of ordered element, which as well contain names.
+The vertices of the topology can be abstracted into two different types, first the bolt and second the spout.
+\textbf{Spout} is the source of tuples into the topology whereas \textbf{Bolt} is a point in the topology where the tuples are transformed.
+A bolt typically does a simple transformation, in order to obtain a complex transformation, chaining of bolts can be done.
 
-\paragraph{Spouts} are the entry point for any data in the Storm Topology.
-As shown in the figure \ref{fig:topo} spouts take data from external source and emits them into the topology as tuples.
-The input source could be a Heterogenous data source.
-
-\paragraph{Bolts} are the processing entities of a Topology.
-It receives raw tuples emitted by spouts or processed tuples sent by another bolt and sends it down the pipe after it has completed its processing.
+The figure \ref{fig:topo} shows a basic Storm Topology with a single spout and three different bolts.
+The spout reads data from external sources, either homogenous or heterogenous and passes on to the bolts, according to the figure Bolt 1 and Bolt 2.
+Bolt 1 and Bolt 2 perform simple transformations and send the results to Bolt 3, which finalizes the data transformation and sends it to the Target.
+A topology as can be noted in the figure might have multiple instances of Bolts or Spouts running, which allows the scalability of the total system.
 
@@
 
+\paragraph{Multiple instances} of the same bolt is available within a topology, thus, Storm allows one to partition the data stream to the bolts, termed as Stream Grouping.
+There are eight available built-in stream groupings which can be used and custom extensions can be made according to the needs of the system.\footnote{\url{http://storm.apache.org/releases/1.1.0/Concepts.html}} The common built-in groupings are:
+
+\begin{enumerate}
+\item{Shuffle Grouping} : Random distribution of the tuples among the target nodes. Load balancing can be achieved using this grouping.
+\item{Fields Grouping} : The tuples are partitioned by the value of a field.
+All the tuples with the same value of a specific field are routed to the same instance of the bolt.
+Can be used if a state of the bolt depends on the value of a field within the tuple.
+\item{All Grouping} : The tuples are replicated to every instance of the bolt.
+\item{Global Grouping} : The tuples are sent only to a single instance of the bolt.
+\end{enumerate}
+
+Each tuple is produced and consumed by bolts, but actual communication happens between the Worker processes.
+Each worker process has a Send thread and a Receive Thread which is responsible for handling the communication.
+As seen in figure \ref{fig:communication}, the Worker Receive thread receives messages from the global buffer and demultiplexes it to the respective tasks running in the User thread pool.
+Each User thread contains two message queues associated with it, the in-queue where incoming data is stored and the out-queue where the outgoing data is stored.
+The user thread, takes messages from the queue, processes them and sends them to the out queue.
+The data in the out queue is processed by the send queue and then passed on to the Worker Send Thread to be communicated into the network.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{topo.png}
+    \caption{Generic Storm Topology}
+    \label{fig:communication}
+   \end{center}
+\end{figure}
+
 \subsection{Storm Features}
+Apache Storm has many features, some of them have been described in this section.
+\subsubsection{Scalable}
+\paragraph{Apache Storm} is a distributed system, and thus highly scalable.
+Adding of worker nodes in a storm is possible to increase the computational capacity of the total system.
+The concept of bolts with limited computations, allows one to increase the number of bolts performing a single task.
+For example, in a scenario where Bolt 1 parses text and Bolt 2 just counts occurrances, it is clear that Bolt 1 takes more time to process data rather than Bolt 2.
+So, increasing the number of Bolt 1 could be a better option in terms of scalability.
+
+\subsubsection{Simple API}
+\paragraph{Storm} provides a simple API for building on top of the pre-defined abstractions.
+Listing \ref{lst:spout} shows a bare bone implementation of a Spout.
+The Spout class extends the BaseRichSpout and implements only four methods, namely, declareOutputFields to declare the output fields this spout will be producing, open to handle initialization phase and close to handle the destroy phase and nextTuple to generate the next tuple.
+The API is similarly simple for implementing a Bolt.
+The complete detail of the API is out of the scope of this report and can be obtained from the official Storm documentation
+
+\begin{lstlisting}[float,caption=Bare Spout Implementation,label=lst:spout]
+public class TwitterFeedSpout extends BaseRichSpout {
+	@Override
+	public void declareOutputFields(OutputFieldsDeclarer declarer) {
+	}
+
+	@Override
+	public void open(Map conf, TopologyContext context, 
+	SpoutOutputCollector collector) {
+	}
+
+	@Override
+	public void nextTuple() {
+	}
+
+	@Override
+	public void close() {
+	}
+}
+\end{lstlisting}
+
+Another feature of the Storm API is that it supports multiple languages using JSON based protocol over Standard IO\footnote{\url{http://storm.apache.org/about/multi-language.html}}.
+This enables code reusability in most cases and makes it easier to port code into storm.
+
 \subsubsection{Fault Tolerance and Resilience}
 \label{sec:resilience}
-Apache storm offers fault tolerance and resilience mechanisms with out of the box support and is tolerant against different types of failures.
+\paragraph{Apache storm} offers fault tolerance and resilience mechanisms with out of the box support and is tolerant against different types of failures.
 Nimbus, being the main master node, is in-charge of all the other worker nodes.
@@
 
+Apache Storm supports two types of processing schemes, \textbf{at least once} where each tuple is processed as the name suggests, and the other \textbf{at most once}, where the tuples are either processed once or dropped in cases of failure.
+To provide \textbf{at least once} semantics, Storm maintains an acker bolt, which keeps track of the tuple flow in the directed acyclic graph.
+
 \begin{table}
@@
 
-Apache Storm supports two types of processing schemes, \textbf{at least once} where each tuple is processed as the name suggests, and the other \textbf{at most once}, where the tuples are either processed once or dropped in cases of failure.
-To provide \textbf{at least once} semantics, Storm maintains an acker bolt, which keeps track of the tuple flow in the directed acyclic graph.
-
 \begin{figure}
@@
 It keeps track of the tuple flow starting from the point where it has been injected into the topology till it leaves the topology.
-After the tuple has left the topology, an \textbf{ack} is sent to the Spout, at this point, it can decide to remove the Tuple from memory.
-
-\section{Stream Processing: Design Considerations}
-A stream is a continuous flow of data from one point to the other.
-A Stream Processing system works on these data streams, the work may just be simple transformations, but could also be as complex as predictive analysis.
-A stream processing system depending on the nature of application has different requirements.
-
-Let us consider the news feed service provided by any social networking site, we know that in any of these websites we can read posts created by our friends or family and it does not matter if we see them at the very instance they post.
-But, there are cases such as a stream of purchases, where we want the data to be visible as soon as available separate design considerations should be made.
-This section focuses on the various considerations that are to be made while designing a stream processing system, specifically using Apache Storm.
-
-\subsection{System Architecture}
-As discussed above, a data processing stream has different requirements and according to these various requirements, the System can be modified to suit these needs.
-The major two System architectures currently available are the Lambda Architecture and the Lambda-less (or Kappa) Architecture.
-
-\subsubsection{Lambda-less Architecture}
-\paragraph{Kappa Architecture} or the lambda-less Architecture is a System Architecture concept where there are only two layers, namely Speed Layer and Serving Layer.The data input to the system is processed at real time to retrieve the results at real time.
-The major advantage of this type of system is that there is minimum latency.
-This type of architecture is more suitable in cases where an accurate view of the data is preferable without a lot of delay.
-Cash transactions might be one of these cases where a minimum delay for transaction visualization is suitable added the calculations done on these data sets are properly formulated as well.
-
-This architecture can be considered has an advantage over the lambda architecture (described later) , that a single code base is to be maintained throughout the life of the software.
-But, any apparent change to the code base results in a need for a duplicate cluster to analyze the data again and update the view.
-A working example of this architecture is presented in the section \ref{sec:usecase}.
-One may use this architecture over the lambda architecture when both the algorithms for the batch processing and stream processing are the same and the requirements mentioned above are the major focus.
-
-\subsubsection{Lambda Architecture}
-\paragraph{Lambda Architecture} is a data processing Architecture designed to handle high volume data utilizing the best of both batch and stream processing.
-Figure \ref{fig:lambda} shows a generic architecture principle for Lambda Architecture.
-The three composed layers are: Speed Layer, Batch Layer and Serving Layer.
-The batch processing layer provides an accurate view of the data but provides results in batches, whereas the speed layer provides current data visualizations which may or may not be accurate depending upon the algorithm.
-The serving layer utilizes both the speed layer as well as the batch layer to fulfill the request to a query.
-
-Lambda architecture requires maintanence of two separate systems as well as code bases, and thus may result in higher cost.
-But, the added advantage is that the batch processing layer code can be optmized for the given data set to obtain better results.
-An example scenario could be a specialized news feed which requires machine learning algorithms to customize the user experience.
-The Speed layer could be used to deliver recent news, but the batch layer could use machine learning to produce the customized feed.
-An added advantage is that Lambda architecture holds the input data unchanged.\cite{lambdastorm}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=.7\textwidth]{lambda.png}
-    \caption{Lambda Architecture}
-    \label{fig:lambda}
-   \end{center}
-\end{figure}
-
-\subsection{Topology Design Considerations}
-A storm topology design with the spouts and bolts and affects the overall system thus should be properly considered.
-A spout should inject data into the topology from external sources whereas the bolt should be a smallest implementable task.
-A storm topology should also be considered with various considerations in mind:
-
-\subsubsection{Scalability:} A designed topology should be considered with scalability in mind.
-The major question is what configuration of bolts enables us to process greater amount of data at real time.
-Let us consider a simple example of word count topology which counts the repetition of words and ranks the most used words.
-It can be performed by simply including a spout and two bolts, the first of which to count the number of words and the second to aggregate and rank it (see fig.
-\ref{fig:ranking} Topology Setup 1).
-Since we should have only one instance of Aggregator, to be able to get the aggregated result, in cases of increased data load, the number of rolling count bolts can be increased, but the Aggregation Bolt cannot be increases thus adding increased load at this node.
-To relax the demand on this node, we can introduce another bolt, the Intermediate Ranking Bolt, which is responsible for ranking the words at certain predefined intervals and passes the intermediate ranks to the aggregator.
-This reduces load on the Aggregator since we can increase the intermediate ranking bolts to cope with the increased load.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\textwidth]{ranking.png}
-    \caption{Possible Topologies for Word Count and Ranking}
-    \label{fig:ranking}
-   \end{center}
-\end{figure}
-
-\subsubsection{Resilience:} Apache storm includes APIs which allow users to develop solutions which are resilient in nature.
-Using the acker bolt as shown in figure \ref{fig:acker}, users can track the progress of the tuple injected into the system.
-Thus, writing a spout which can replay the tuple in case of failures allows us to guarantee the processing of data.
-For additional details please follow Section \ref{sec:resilience}.
-
-\subsubsection{Interoperability:} The designed topology should be interoperable with different sources of data, and by default should work out of the box by adding simply other Spouts in the topology.
-This also amounts to the maintainability of the code.
+After the tuple has left the topology, an \textbf{ack} is sent to the Spout, at this point, it removes the entry from memory.
 
@@
 \paragraph{This} is not the only possible solution, there are multiple other tools such as \textbf{Spark Streaming}, \textbf{Samza}, etc.
-but this paper concentrates Apache Storm as it is in the spotlight of this paper, for further details regarding other available tools and their comparisons with Apache Storm, please consider the section \ref{sec:othertools} for further details.
+but this paper concentrates Apache Storm as it is in the spotlight of this paper, for further details regarding other available tools and their comparisons with Apache Storm, please consider the section \ref{sec:toolcomparision} for further details.
 
@@
 The only spout available in this example uses Twitter Hosebird Client (\url{https://github.com/twitter/hbc}) to connect to Twitter Streaming Server and consume the Tweets generated in real time.
-Whereas the Bolts were programmed in java and the code can be viewed in \url{https://github.com/hakku276/storm-sentiment-analysis}.
+Whereas the Bolts were programmed in java and the complete code can be viewed in \url{https://github.com/hakku276/storm-sentiment-analysis}.
+A short extract how the topology was created can be seen in listing  \ref{lst:toposetup}.
+We can see that a spout has been declared in line 2, then a parser bolt and finally a aggregator bolt with parallelism 1.
+It is then submitted to storm (nimbus).
+
+\begin{lstlisting}[float,caption=Bare Spout Implementation,label=lst:toposetup]
+TopologyBuilder builder = new TopologyBuilder();
+builder.setSpout("spout", new TwitterFeedSpout(tags));
+builder.setBolt("parse", new TweetParseBolt())
+	.shuffleGrouping("spout");
+builder.setBolt("aggregation", new TwitterAggregationBolt(), 1)
+	.shuffleGrouping("parse");
+StormSubmitter.submitTopology("SentimentAnalyzer", config, 
+	builder.createTopology());
+\end{lstlisting}
 
@@
 
-\section{Other Tools}
- \label{sec:othertools}
+\section{Comparision to Related Work}
+ \label{sec:toolcomparision}
 Stream processing is a topic under constant focus at present and a lot of tools as well as models have been designed and created to support it.
 This section makes a comparison of the available widely accepted tools today.
-The available tools are mentioned below in brief such as:
 
@@
  \end{enumerate}
-
- \subsection{Comparison between Storm, Spark and Samza}
-This sub section compares the three most widely used tools for Stream Processing, i.e. Storm, Spark and Samza.
 
@@
 The fault tolerance model for Storm has been extensively described in section \ref{sec:resilience}.
+In cases of worker failure, as described earlier, Storm restarts the worker and if a tuple that is being processed by this node was lost, is again replayed.
 Similarly, Spark Streaming and Apache Samza also use various techniques to resolve issues while deployed.
-Spark Streaming uses write ahead logs(journals) to keep track of all the data operation, and during failure of a worker, the worker is restarted and a state restore is done by reading the logs.
-This ensures that the data is processed as required.
-In comparison to Apache Storm, Storm continues to process incoming tuples even when the master node fails, but in Spark, the worker nodes stop working when the master node fails until a new leader starts.
-This is an added bonus to Storm, since a failed master manual restart would reinstate the state of the system without many changes.
+Spark Streaming uses parallel worker recovery which is possible with the help of RDDs and Apache Samza uses checkpointing to restore a failed state.
+Due to the parallel recovery of Spark Streaming, in my opinion it proves better in worker failure resilience, since Storm can recognize that a node has failed and can restart it, but it cannot know a tuple has been lost until the tuple has timed out.
+Time outs take extra time and results in slower processing.
+
+In case of master node failure, Storm worker node continues to process incoming tuples, but in Spark, the worker nodes stop working when the master node fails until a new leader starts.
+This is an added bonus to Storm, since a failed master manual restart would reinstate the state of the system without changes.
+This is a better resilience feature than provided by Spark Streaming.
 
@@
 \begin{center}
-\includegraphics[width=.75 \textwidth]{comparisiongraph.png}
+\includegraphics[width=0.85\textwidth]{comparisiongraph.png}
 \caption{Performance Comparision between Apache Storm, Apache Flink and Spark Streaming \cite{benchmark}}
@@
 The tool is comparatively easy to setup and configure, with out of the box configurations which is production ready.
-The storm API is relatively simple and does not have a complicated method calls, the programming is based on event driven programming and since storm supports read and writes based on standard IO, utilizing previously written code could be relatively easier than in the other available tools.
+The API is as well relatively simple and does not have a complicated method calls, the programming is based on event driven programming and since storm supports read and writes based on standard IO, utilizing previously written code could be relatively easier than in the other available tools.
 Data is currently being generated and consumed at an unprecedented rate, thus stream processing technologies is a requirement and is gaining popularity and attention.
 With all the various tools available, one can decide on which tool best suites ones needs.
-According to the benchmarks, if minimum latency is required, Apache Storm or Apache Flink or any system with Continuous data flow model should be preferred.
+According to the benchmarks, if minimum latency is required, Apache Storm or Apache Flink or any system.
 Whereas if high throughput is required with a non strict latency criteria, Spark streaming could prove to be a better option.

### 2018-02-20 paper.tex
Finalized the report

@@
 Apache Storm is one of the available open source tools for Stream Data Processing.
-This document provides a brief overview of the intricacies of the Architecture of Storm and tends to proceed further on the details of considerations required to design a Solution based on Storm.
-The design considerations relating to deployment of Storm System as a Lambda or a Lambda-less Architecture and the sketch of the various internal components of Storm i.e.
-Bolts, Spouts.
+This document explores the Apache Storm architecture in detail and narrates the most important features.
+Resilience to faults being an important factor in Stream processing due the required availability.
 We consider a practical implementation of Storm to try and express the theoretical design considerations in practice.
-Finally we list out the other available tools and compare them to Storm with regard to Performance, Resilience and various other parameters.
+We also introduce in brief tools which are currently used in general, namely Apache Samza, Apache Spark Streaming and Apache Flimk.
+Compare them to Storm with regard to Performance, Resilience and various other parameters.
 \end{abstract}
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{example.png}
+    \includegraphics[width=0.85\textwidth]{example.png}
     \caption{An Example Stream Processing Scenario}
@@
   \begin{center}
-    \includegraphics[width=.7\textwidth]{arch.png}
+    \includegraphics[width=0.85\textwidth]{arch.png}
     \caption{Storm Physical Architecture}
@@
   \begin{center}
-    \includegraphics[width=.7\textwidth]{worker.png}
+    \includegraphics[width=.85\textwidth]{worker.png}
     \caption{Worker Node Architecture}
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{topo.png}
-    \caption{Generic Storm Topology}
+    \includegraphics[width=\textwidth]{communication.png}
+    \caption{Message flow in a Worker}
     \label{fig:communication}
@@
   \begin{center}
-    \includegraphics[width=.7\textwidth]{topoimpl.png}
+    \includegraphics[width=\textwidth]{topoimpl.png}
     \caption{Topology Implementation for Sentiment Analysis}
